---
title: "Site occupancy modelling of Brown-headed Nuthatch Ebird data"
data: today
author: Max Savery
format: html
---

```{r set-up, include=FALSE, message=FALSE}
library(auk)
library(terra)
library(tidyterra)
library(tidyverse)
library(rnaturalearth)
library(sf)
library(lubridate)
library(dggridR)
library(unmarked)
library(exactextractr)
library(AICcmodavg)
library(stringr)

select <- dplyr::select

base_data_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\data\\"
ebd_download_dir <- "ebd_US_bnhnut_201901_201912_smp_relJul-2023"
# Bool for doing inital auk processing
auk_process <- F
occ_fit <- T
map_proj <- st_crs("ESRI:102003")
```

In this report, we discuss site-occupancy modelling of the Brown-headed Nuthatch in Tennessee. Nuthatch data is first downloaded from the Ebird site https://science.ebird.org/en/use-ebird-data/download-ebird-data-products. You must first initally ask for access to the Ebird data. 

The format for this analysis closely follows the code and advice from https://cornelllabofornithology.github.io/ebird-best-practices/occupancy.html#occupancy-predict.

The Nuthatch data was first filtered using the Ebird data service, for the period of January 1 to December 31, 2019. Brown-headed nuthatches are non-migratory, so we use the full year of 2019. With this dataset, we proceeded to further process it using the R package auk. These steps include verifying the correct time period is filtered, selecting complete checklists (those with all species reported), and only those with stationary and travelling protocols (excluding historical and incidental). For pure PO data modelling we can use the incidental observations as well, but for this occupancy analysis we'll stick to the 2 protocols.

Here only complete checklists are kept, which means that even if we keep the incidental they will be dropped as they are not considered complete. They lack information like time and duration, and distance traveled. For more about the types of protocols: https://support.ebird.org/en/support/solutions/articles/48000950859-guide-to-ebird-protocols
```{r ebird-loading, eval=auk_process}
ebd_nh <- auk_ebd(file.path(base_data_dir, ebd_download_dir, "ebd_US_bnhnut_201901_201912_smp_relJul-2023.txt"),
                  file_sampling = file.path(base_data_dir, ebd_download_dir, "ebd_US_bnhnut_201901_201912_smp_relJul-2023_sampling.txt"))
ebd_nh %>% auk_date(date = c("2019-01-01", "2019-12-31")) %>%  auk_protocol(protocol = c("Stationary", "Traveling")) %>% auk_complete() -> ebd_nh_filtered
auk_filter(ebd_nh_filtered, file = file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_po_2019.txt"), 
                            file_sampling=file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_2019_sampling.txt"), overwrite=T) 
nuthatch_obs <- read_ebd(file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_po_2019.txt"))
# TODO: Make plots only using the PO data and not the merged sampling + PO data
nuthatch_sampling <- read_sampling(file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_2019_sampling.txt"))
```
We then use the zerofill function to combine the Presence-Only observations with the sampling data, in order to generate Presence-Absence type data.
```{r zerofill, eval=auk_process}
nuthatch_zf <- auk_zerofill(nuthatch_obs, nuthatch_sampling, collapse = TRUE)
# Some 4 million rows
head(nuthatch_zf)
```
There are a few other processing steps outlined in the tutorial, in order to properly filter the PA data

One important processing step is based off the following statement from the guide:
  "Based on our experience working with these data, we suggest restricting checklists to less than 5 hours long and 5 km in length, and with 10 or fewer observers." This filtering is reflected below. In eBird a single checklist can be shared between multiple people, so limiting this to less than 10 removes some variability in the observation gathering. Later on, this will be filtered to less than 5 observers, but it is left at 10 for data exploration.
  
Time is coerced to a decimal value between 0-24, and distance travelled is set to 0 for stationary protocols. Finally, "X" values in the count data are converted to NA.
```{r filter-zf, eval=auk_process}
time_to_decimal <- function(x) {
  x <- hms(x, quiet = TRUE)
  hour(x) + minute(x) / 60 + second(x) / 3600
}

# clean up variables
nuthatch_zf_filter1 <- nuthatch_zf %>% 
  mutate(
    # convert X to NA
    observation_count = if_else(observation_count == "X", 
                                NA_character_, observation_count),
    observation_count = as.integer(observation_count),
    # effort_distance_km to 0 for non-travelling counts
    effort_distance_km = if_else(protocol_type != "Traveling", 
                                 0, effort_distance_km),
    # convert time to decimal hours since midnight
    time_observations_started = time_to_decimal(time_observations_started),
    # split date into year and day of year
    year = year(observation_date),
    day_of_year = yday(observation_date)
  )

# additional filtering
nuthatch_zf_filter2 <- nuthatch_zf_filter1 %>% 
  filter(
    # effort filters
    duration_minutes <= 5 * 60,
    effort_distance_km <= 5,
    # 10 or fewer observers
    number_observers <= 10)

nuthatch <- nuthatch_zf_filter2 %>% 
  select(checklist_id, observer_id, sampling_event_identifier,
         scientific_name,
         observation_count, species_observed, 
         state_code, locality_id, latitude, longitude,
         protocol_type, all_species_reported,
         observation_date, year, day_of_year,
         time_observations_started, 
         duration_minutes, effort_distance_km,
         number_observers)
# Save that csv for later use
write_csv(nuthatch, file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_for_occ.csv"), na = "")
```
```{r read_ebd}
# Read pre-processed data
nuthatch_obs <- read_ebd(file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_po_2019.txt"))
nuthatch <- read_csv(file.path(base_data_dir, ebd_download_dir, "nuthatch_filtered_for_occ.csv")) %>% 
  mutate(year = year(observation_date),
         # occupancy modeling requires an integer response
       species_observed = as.integer(species_observed))
```
Once we have the nuthatch data processed, we can start to consider the ecological covariates that will be useful for modelling occupancy. 
We will focus on the Brown-headed nuthatch in Tennessee, so the first step is to load the state boundary polygon.
```{r state-bound}
state_bound <- read_sf(file.path(base_data_dir, "us_states/GOVTUNIT_Tennessee_State_GPKG/GOVTUNIT_Tennessee_State_GPKG.gpkg")) %>% 
  st_transform(crs = map_proj) %>% 
  st_geometry()

# Convert to sf for spatial processing
nuthatch_sf <- nuthatch %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% st_transform(crs="ESRI:102003") %>% st_geometry()
st_crs(nuthatch_sf) == st_crs(state_bound) 

nuthatch_po_sf <- nuthatch_obs %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% st_transform(crs="ESRI:102003") %>% st_geometry()
st_crs(nuthatch_po_sf) == st_crs(state_bound) 

# Get the points within the state
state_pp <- st_intersection(nuthatch_sf, state_bound)
state_pp_po <- st_intersection(nuthatch_po_sf, state_bound)

# More advanced using tidyterra and ggplot
p <- ggplot() + 
  #geom_spatvector(data=state_grid, fill = 'transparent', colour="lightblue") +
  geom_sf(data = state_bound, color=alpha("white",0.9)) + 
  geom_sf(data=state_pp, color=alpha("darkgrey",0.2), size=0.5)+
  geom_sf(data=state_pp_po, color=alpha("#FFC81C",0.5), size=0.5)+
  ggtitle("Brown-headed Nuthatch checklists in Tennessee") +
  theme_minimal() +
  theme(text=element_text(size=10)) +
  coord_sf()
print(p)
#fig_name="data/ebird/ebd_checklist_nuthatch.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
```
We can see the nuthatch observations within Tennessee as well as all the sites at which absences of nuthatches were recorded. 

Next, let's prepare the covariate data. This comes in the form of raster data, particularly PROBA-V landcover and Modis landcover dynamics product. 

To download the PROBA-V satellite data, go to https://land.copernicus.eu/global/products/lc and use the world viewer. Select the appropriate section/s 
of the globe for your use-case. On this case, we simply need one section that appropriately covers Tennessee. Information regarding the product itself can be found in the documentation: https://land.copernicus.eu/global/sites/cgls.vito.be/files/products/CGLOPS1_PUM_LC100m-V3_I3.4.pdf. Important for this analysis are the classifications of each landcover type.

TODO: Include picture of this from the PDF.
TODO: Save processed rasters. Either check if files exist or provide switch
```{r lc-process}
landcover_filename <- file.path(base_data_dir, "copernicus_landcover/W100N40_PROBAV_LC100_global_v3.0.1_2019-nrt_Discrete-Classification-map_EPSG-4326.tif")
file.exists(landcover_filename)
lc_se_us <- rast(landcover_filename) 
# Project state to raster crs
# Then crop raster in crs of raster, and convert raster back to original state crs
prj_state <- terra::project(vect(state_bound), lc_se_us)
crop_lc_filename <- file.path(base_data_dir, "copernicus_landcover/cropped_lc.tif")
if (!file.exists(crop_lc_filename)){
  crop_lc_rast <- crop(lc_se_us, prj_state)
  writeRaster(crop_lc_rast, crop_lc_filename)
}else{
  crop_lc_rast <- rast(crop_lc_filename)
}
state_pp_prj <- terra::project(vect(state_pp), lc_se_us)
state_pp_po_prj <- terra::project(vect(state_pp_po), lc_se_us)

# Plot landcover
p <- ggplot() + 
  geom_spatraster(data=crop_lc_rast) +
  geom_sf(data=state_pp_prj, color=alpha("darkgrey", 0.3))+
  geom_sf(data=state_pp_po_prj, color=alpha("#FFC81C",0.5), size=0.5)+
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch and landcover in Tennessee")
print(p)
#fig_name="data/ebird/landcover.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
```
We also include information regarding the vegetation index. We use the Enhanced Vegetation Index MODIS product. Different layers from this download are described here: https://lpdaac.usgs.gov/products/mcd12q2v061/
Additionally, https://www.usgs.gov/landsat-missions/landsat-enhanced-vegetation-index#:~:text=EVI%20is%20similar%20to%20Normalized,in%20areas%20with%20dense%20vegetation provides further information regarding the index.

To download the data, the service at https://appeears.earthdatacloud.nasa.gov/task/area allows you to upload a polygon for the region of interest and will subsequently provide
a link to download the corresponding MODIS product.
```{r modis-vi}
modis_evi_filename <- file.path(base_data_dir, "modis_landcover_dynamics/MCD12Q2.061_EVI_Area_0_doy2019001_aid0001.tif")
file.exists(modis_evi_filename)
modis_evi_rast <- rast(modis_evi_filename) 
# Project full raster then crop.
crop_evi_filename <- file.path(base_data_dir, "modis_landcover_dynamics/cropped_evi.tif")
if(!file.exists(crop_evi_filename)){
  evi_rast_prj <- terra::project(modis_evi_rast, lc_se_us)
  crop_evi_rast <- crop(evi_rast_prj, prj_state)# Plot evi
  writeRaster(crop_evi_rast, crop_evi_filename)
}else{
  crop_evi_rast <- rast(crop_evi_filename)
}

p <- ggplot() + 
  geom_spatraster(data=crop_evi_rast) +
  geom_sf(data=state_pp_prj, color=alpha("darkgrey", 0.3))+
  geom_sf(data=state_pp_po_prj, color=alpha("#FFC81C",0.5), size=0.5)+
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch and EVI in Tennessee")
print(p)
#fig_name="data/ebird/evi.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
```
Now we need to process the landcover data so that it is  more suitable for occupancy modelling. As per the eBird guide to occupancy modelling
convert the landcover classes to proportions of the categories within a specified buffer.
According to the tutorial, "approximately 2.5 km by 2.5 km neighborhood (5 by 5 MODIS cells) centered on the checklist location is 
sufficient to account for the spatial precision in the data when the maximum distance of travelling counts has been limited to 5 km". 
We proceed according to that recommendation.

The code block will only run here if the covariate file has not already been generated.
```{r covar-processing}
neighborhood_radius <- 5000 * ceiling(max(res(crop_lc_rast))) / 2
neighborhood_radius
nuthatch_unique <- nuthatch %>% distinct(locality_id, latitude, longitude) 
# Select only the unique observations within the state
nuthatch_unique_state <- nuthatch_unique %>% st_as_sf(coords = c("longitude", "latitude"), crs=4326) %>% 
  # transform to state crs projection
  st_transform(crs = crs(state_bound)) %>% 
  st_intersection(state_bound)
nuthatch_unique_state

nuthatch_buffer <- nuthatch_unique_state %>% 
  # buffer to create neighborhood around each point
  st_buffer(dist = neighborhood_radius)
nuthatch_buffer

# Take a look at the buff on -headed Nuthatch  in Tennessee the map
# The buffer will have all sampled sites, not just observed PO sites
buff_prj <- st_transform(nuthatch_buffer, crs=crs(crop_lc_rast))
crs(buff_prj)
p <- ggplot() + 
  geom_spatraster(data=crop_lc_rast) +
  geom_sf(data=buff_prj, color=alpha("black", 1), linewidth=0.5)+
  geom_sf(data=state_pp_prj, color=alpha("orange", 0.5))+
  #geom_sf(data = state_bound, color=alpha("white",0.9), fill='transparent') + 
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Buffer on observed sites")
print(p)
#fig_name="data/ebird/site_buffer.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")`

# Create the fractional covariates
lc_ext_frac <- exact_extract(crop_lc_rast, buff_prj, "frac", progress=F)
lc_ext_frac
dim(lc_ext_frac)
max_cols <- max.col(lc_ext_frac)
max_col_names <- colnames(lc_ext_frac)[max_cols]
mode_sites <- as.numeric(str_sub(max_col_names, 6, -1))
# Take mode
lc_ext_mode <- exact_extract(crop_lc_rast, buff_prj, "mode", progress=F)
# Manual mode from frac should equal mode using function .
stopifnot(lc_ext_mode == mode_sites)

# Let's do the same for the EVI data
evi_buff <- exact_extract(crop_evi_rast, buff_prj, "mean")
evi_df <- data.frame(evi=evi_buff, locality_id=buff_prj$locality_id)

# Mapping unique sites to covariates using the locality_id
dim(nuthatch_unique_state)
lc_ext_frac$locality_id <- buff_prj$locality_id
names(lc_ext_frac)
#nuthatch_unique_state$landcover <- lc_ext_mode 
# In this way we will keep the buffer polygons
nuthatch_unique_covars <- inner_join(buff_prj, lc_ext_frac, by=c("locality_id"))
nuthatch_unique_covars <- inner_join(nuthatch_unique_covars, evi_df, by=c("locality_id"))
dim(nuthatch_unique_covars)
nuthatch_unique_covars
names(nuthatch_unique_covars)
class(nuthatch_unique_covars)
  
evi_rast <- terra::rasterize(vect(nuthatch_unique_covars), crop_evi_rast, field="evi")
evi_rast
evi_rast[is.na(evi_rast)] <- 0
plot(evi_rast)

p <- ggplot() + 
  geom_sf(data=buff_prj, color=alpha("black", 1), linewidth=0.5)+
  geom_sf(data = prj_state, color=alpha("black",0.5), fill='transparent', linewidth=0.7) + 
  geom_spatraster(data=evi_rast) +
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7, na.value="white") +
  theme_minimal()+
  ggtitle("Nuthatch EVI in Tennessee at Nuthatch sites")
print(p)
#fig_name="data/ebird/site_evi.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
```
Having processed the covariate data for use in the occupancy model, we now take the final processing steps. We 

The Cornell Labs all about birds site describes the environment of the Nuthatch as follows:
(https://www.allaboutbirds.org/guide/Brown-headed_Nuthatch/id#:~:text=Brown%2Dheaded%20Nuthatches%20occur%20in,in%20mixed%20pine%2Dhardwood%20stands.)

"Brown-headed Nuthatches occur in Southeastern pine forests in areas with loblolly, shortleaf, longleaf, and slash pine trees. They are most common in open and mature pine forests and are less common in mixed pine-hardwood stands."

Based on this description, we select a subset of the landcover classes to use within the occupancy models. We use the variables (with classificaiton ids)

30: Herbaceous vegetation
40: Agriculture
111: Closed forest, evergreen needle leaf
114: Closed forest, deciduous broad leaf
115: Closed forest, mixed
116: Closed forest, unknown
121: Open forest, evergeen needle leaf
126: Open forest, unknown

There are more specific occupancy filtering steps taken here, to ensure that the data follows the PA assumptions. The filter_repeat_visits function does this. By setting annual_closure=False, we consider a year the period of occupancy closure. Furthermore, only sites are selected for occupancy modelling if they have been visited at least twice and less than 10 times, sampling out some sampling bias. If they are visited more than 10 times, 10 checklists are randomly sampled from these visits. If we want to define periods of closure less than a year, there is the argument n_days that let's us do this. 

Once we run this function, the n_observations column (as well as closure_id and site) is added to the dataset, which counts the total number of observations over all the visits. We will also get a column for each visit, as per the occupancy-type likelihood assumptions.
```{r occ-processing, eval=TRUE}
# Then final step to prepare occupancy data for modelling
# combine ebird and modis data
# This will reduce nuthatch to only locality_id in the state
ebird_habitat <- inner_join(nuthatch, nuthatch_unique_covars, by = c("locality_id"))
# Then filter based on a few criteria I need to check
# so that the data are closed PO data
ebird_habitat$number_observers
ebird_filtered <- filter(ebird_habitat, 
                         number_observers <= 5)

occ <- filter_repeat_visits(ebird_filtered, 
                            min_obs = 2, max_obs = 10,
                            annual_closure = TRUE,
                            date_var = "observation_date",
                            site_vars = c("locality_id", "observer_id"))

class(occ)
colnames(occ)
covar_names <- c(colnames(occ)[26:length(colnames(occ))], "latitude", "longitude")
covar_names
select_covar <- covar_names[c(1,2, 7, 8, 9, 10, 11, 13, 14, 15, 16)]
select_covar
#covar_labels <- c("herbaceuos_vegetation", "agriculture", "closed_forest_evergreen_needle", 
#                  "close_forest_decid_broad", "closed_forest_mixed", 
#                  "closed_forest_unknown", "open_forest_evergreen_needle", 
#                  "open_forest_decid_needle", "open_forest_unknown", "evi")

# Take a look at all the covariates by plotting rasterized versions of them, which is faster
covar_names[1:(length(covar_names)-3)]
for(covar in covar_names[1:(length(covar_names)-3)]){
  covar_rast <- terra::rasterize(vect(nuthatch_unique_covars), crop_evi_rast, field=covar)
  p <- ggplot() + 
    geom_spatraster(data=covar_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle(paste(covar, " in Tennessee at Nuthatch sites", sep=""))
  print(p)
  #fig_name <- paste("data/ebird/", covar, "_covar_tenn.png", sep="")
  #ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
}
```
We can take a look at all the covariate surfaces. Based on this and the known habitat of the Brown-headed nuthatch, we select the 
covariates for occupancy and detection.
```{r occ_df}
# Use closed forests for detection prob
occ$protocol_type <- as.factor(occ$protocol_type)
occ_wide <- format_unmarked_occu(occ, 
                                 site_id = "site", 
                                 response = "species_observed",
                                 site_covs = select_covar,
                                 obs_covs = c("duration_minutes", "number_observers",  "effort_distance_km",
                                              "time_observations_started", "protocol_type", 
                                              "frac_111", "frac_114", "frac_116", "evi"))

```
There is an additional step taken to reduce spatial bias. This is creating a grid over the region and sampling sites within that. In doing so, we reduce the number of sites surveyed at the most convenient locations. We can view the generated grid as well.
```{r spatial-smapling}
# Do some spatial subsampling to reduce bias
dggs <- dgconstruct(spacing = 5)
# get hexagonal cell id for each site
occ_wide_cell <- occ_wide %>% 
  mutate(cell = dgGEO_to_SEQNUM(dggs, longitude, latitude)$seqnum)
# sample one site per grid cell
occ_ss <- occ_wide_cell %>% 
  group_by(cell) %>% 
  sample_n(size = 1) %>% 
  ungroup() %>% 
  select(-cell)
# calculate the percent decrease in the number of sites
1 - nrow(occ_ss) / nrow(occ_wide)
occ_ss

hexagons <- dgcellstogrid(dggs, unique(occ_wide_cell$cell)) %>% 
  st_as_sf()

# View the grid
p <- ggplot() +
  geom_sf(data = hexagons) +
  geom_sf(data=state_pp_prj, color=alpha("orange", 0.1))+
  theme_bw()
print(p)

# Unmarked formatting
occ_um <- formatWide(occ_ss, type = "unmarkedFrameOccu")
summary(occ_um)
```
We can finally fit the occupancy model.
```{r occ-model, eval=occ_fit}
occ_model <- occu(~ duration_minutes + frac_111 + frac_114 + frac_116 +
                    number_observers + effort_distance_km + time_observations_started + protocol_type
                  ~ frac_30 + frac_40 + frac_111 + frac_114 + frac_115 + frac_116 + frac_121 +
                    frac_126, 
                  data = occ_um)
summary(occ_model)
occ_gof <- mb.gof.test(occ_model, nsim = 10, plot.hist = T)
occ_gof
occ_gof$chisq.table <- NULL
print(occ_gof)
# Just hacking this together based on already existing data
# Using the proportions for the buffers already calculated
occ_pred <- unmarked::predict(occ_model, 
                    newdata = nuthatch_unique_covars,
                    type = "state")
occ_pred$Predicted
nuthatch_unique_covars
nuthatch_unique_covars$preds <- occ_pred$Predicted

# Prepare to convert to raster
preds_select <- nuthatch_unique_covars %>% select(geometry, preds)
preds_select
class(preds_select)
# We can just plot as vector
plot(preds_select)
preds_select$preds
preds_select$geometry

r_pred <- preds_select %>% 
  # convert to spatial features
    #st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  st_transform(crs = crs(crop_lc_rast))

# TODO
# This isn't using the prediction probabilities for some reason
r_pred <- terra::rasterize(r_pred, crop_lc_rast)
r_pred
plot(r_pred)
plot(prj_state, add=T)
```

