---
title: "Site occupancy modelling of Brown-headed Nuthatch Ebird data"
data: today
author: Max Savery
format: pdf
---

```{r set-up}
# Load the ebird data in the way it will be used in the sampling procedure, and fit it to a Bayesian model.
library(sf)
library(terra)
library(tidyterra)
library(sf)
library(auk)
library(tidyverse)
library(lubridate)
library(exactextractr)
library(cmdstanr)
library(bayesplot)


select <- dplyr::select

# Set important global variables 
base_data_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\data\\"
working_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\code\\optimal_design_site_occ"
ebd_download_dir <- "ebd_US_bnhnut_201901_201912_smp_relJul-2023"
lc_path <- "copernicus_landcover/W100N40_PROBAV_LC100_global_v3.0.1_2019-nrt_Discrete-Classification-map_EPSG-4326.tif"
modis_path <- "modis_landcover_dynamics/MCD12Q2.061_EVI_Area_0_doy2019001_aid0001.tif"
elev_path <- "elevation_aster/ASTGTM_NC.003_ASTER_GDEM_DEM_doy2000061_aid0001.tif"
map_prj <- st_crs("ESRI:102003")

# Source modules
source(file.path(working_dir, "data_utils", "load_ebird_data.R"))
source(file.path(working_dir, "experimental_design_functions.R"))
```

```{r load-initial-data}
map_path <- file.path(base_data_dir, "us_states/GOVTUNIT_Tennessee_State_GPKG/GOVTUNIT_Tennessee_State_GPKG.gpkg")
state_bound <- load_map(map_prj, map_path)

# TODO: Include preprocessing for creating the file "nuthatch_filtered_for_occ.csv"
nuthatch_list <- load_ebird(base_data_dir, ebd_download_dir, state_bound, map_prj)
state_pa <- nuthatch_list$state_pa
state_po <- nuthatch_list$state_po

# Plot the data
p <- ggplot() + 
  #geom_spatvector(data=state_grid, fill = 'transparent', colour="lightblue") +
  geom_sf(data = state_bound, color=alpha("white",0.9)) + 
  geom_sf(data=state_pa, color=alpha("darkgrey",0.2), size=0.5)+
  geom_sf(data=state_po, color=alpha("#FFC81C",0.5), size=0.5)+
  ggtitle("Brown-headed Nuthatch checklists in Tennessee") +
  theme_minimal() +
  theme(text=element_text(size=10)) +
  coord_sf()
print(p)

crop_lc_rast <- load_landcover(base_data_dir, lc_path, state_bound)
# Convert crs to the landcover
state_pa_prj <- terra::project(vect(state_pa), crop_lc_rast)
state_po_prj <- terra::project(vect(state_po), crop_lc_rast)
prj_state <- terra::project(vect(state_bound), crop_lc_rast)
state_pa_prj <- st_transform(state_pa, crs(crop_lc_rast))
state_po_prj <- st_transform(state_po, crs(crop_lc_rast))
prj_state <- st_transform(state_bound, crs(crop_lc_rast))


# Plot landcover
p <- ggplot() + 
  geom_spatraster(data=crop_lc_rast) +
  #geom_sf(data=state_pp_prj, color=alpha("darkgrey", 0.3))+
  geom_sf(data=state_po_prj, color=alpha("#FFC81C",0.5), size=0.5)+
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch and landcover in Tennessee")
print(p)
#fig_name="data/ebird/landcover.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")

# Get modis data
crop_evi_rast <- load_evi(base_data_dir, modis_path, prj_state, crop_lc_rast)
p <- ggplot() + 
  geom_spatraster(data=crop_evi_rast) +
  #geom_sf(data=state_pp_prj, color=alpha("darkgrey", 0.3))+
  geom_sf(data=state_po_prj, color=alpha("#FFC81C",0.5), size=0.5)+
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch and EVI in Tennessee")
print(p)

crop_elev_rast <- load_elevation(base_data_dir, elev_path, prj_state)
# Shouldn't need to project either source of rast for elevation rasters
stopifnot(crs(crop_elev_rast)==crs(crop_lc_rast))
stopifnot(st_crs(crop_elev_rast)==st_crs(prj_state))
p <- ggplot() + 
  geom_spatraster(data=crop_elev_rast) +
  #geom_sf(data=state_pp_prj, color=alpha("darkgrey", 0.3))+
  geom_sf(data=state_po, color=alpha("#FFC81C",0.5), size=0.5)+
  geom_sf(data = prj_state, color=alpha("white",0.9), fill='transparent', linewidth=0.4) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch and elevation in Tennessee")
print(p)
```

```{r process-covar}
# Need to use crs for grid with units meters but then transform to the crs of raster, so we also us the pp data in crs in meters
# TODO: Consider transforming raster (nn for categorical;bilinear for continuous)
subgrid <- create_pp_grid(state_bound, crs(crop_lc_rast))
stopifnot(st_crs(subgrid)==st_crs(crop_lc_rast))

p <- ggplot() + 
  geom_spatvector(data=subgrid, color=alpha("#B6B6B6", 0.5)) +
  #geom_spatvector(data=subgrid, aes(fill=counts), color=alpha("#B6B6B6", 0.5)) +
  geom_sf(data=state_po, color=alpha("#FFC81C",0.5), size=0.5)+
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7) +
  theme_minimal()+
  ggtitle("2.5km by 2.5km grid over Tennessee")
print(p)
#fig_name="data/ebird/state_grid.png"
#ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")

pp_counts <- aggregate_point_counts(subgrid, state_po_prj)
pp_counts[which(pp_counts>1)]
hist(pp_counts[which(pp_counts>1)])
subgrid$counts <- pp_counts

# We can plot the subgrid (either as sfc or sf) with spatvector
# make_grid produces sfc but st_sf turns this to sf, to which 
# we can add attributes such as point counts
p <- ggplot() + 
  geom_spatvector(data=subgrid, aes(fill=counts), color=alpha("#B6B6B6", 0.5)) +
  scale_fill_viridis_c(begin=0.2, end=1, option="magma",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch aggregated intensity in Tennessee")
print(p)

# Next calculate covariate aggregation based on the grid.
evi_buff <- exact_extract(crop_evi_rast, subgrid, "mean", progress=T)
plot(evi_buff)
# TODO: Add ebird covariates from the buuffer: number_observers and duration_minutes
agg_covars <- generate_gridded_covariates(subgrid, crop_lc_rast, crop_evi_rast, crop_elev_rast, state_po_prj)
agg_covars$lc_frac
subgrid$evi <- agg_covars$evi_agg
subgrid$elev <- agg_covars$elev_agg
subgrid_covars <- bind_cols(subgrid, agg_covars$lc_frac)

p <- ggplot() + 
  geom_spatvector(data=subgrid_covars, aes(fill=counts), color=alpha("#B6B6B6", 0.5)) +
  scale_fill_viridis_c(begin=0.2, end=1, option="magma",alpha=0.7) +
  theme_minimal()+
  ggtitle("Brown-headed Nuthatch aggregated intensity in Tennessee")
print(p)

# rasterized version of site evi 
evi_rast <- terra::rasterize(subgrid_covars, crop_evi_rast, field="evi")
#evi_rast[is.na(evi_rast)] <- 0
plot(evi_rast)

# rasterized version of site elev 
elev_site_rast <- terra::rasterize(subgrid_covars, crop_evi_rast, field="elev")
elev_site_rast[is.na(elev_site_rast)] <- 0
plot(elev_site_rast)

p <- ggplot() + 
#  geom_sf(data = prj_state, color=alpha("black",0.5), fill='transparent', linewidth=0.7) + 
  geom_spatraster(data=evi_rast) +
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7, na.value="white") +
  theme_minimal()+
  ggtitle("Nuthatch EVI in Tennessee at Nuthatch sites")
print(p)

p <- ggplot() + 
  geom_sf(data = prj_state, color=alpha("black",0.5), fill='transparent', linewidth=0.7) + 
  geom_spatraster(data=elev_site_rast) +
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7, na.value="white") +
  theme_minimal()+
  ggtitle("Elevation in Tennessee at Nuthatch sites")
print(p)

colnames(subgrid_covars)
covar_names <- c(colnames(subgrid_covars)[5:(length(colnames(subgrid_covars))-1)])
covar_names
length(covar_names)
# List of all covar labels
covar_labels <- c("shrubs", "herbaceous_vegetation", "agriculture", "urban", "bare/sparse", "water", "wetland", "closed_forest_evergreen_needle", 
                  "close_forest_decid_broad", "closed_forest_mixed", "closed_forest_unknown", "open_forest_evergreen_needle", 
                  "open_forest_decid_broad leaf", "open_forest_unknown")
covar_labels
names(covar_names) <- covar_labels
covar_names

# Take a look at all the fractional covariates by plotting rasterized versions of them, which is faster
for(i in 1:length(covar_names[1:length(covar_names)])){
  covar <- covar_names[i]
  covar_name <- covar_labels[i]
  covar_rast <- terra::rasterize(vect(subgrid_covars), crop_evi_rast, field=covar)
  p <- ggplot() + 
    geom_spatraster(data=covar_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle(paste(covar, "; ", covar_name, " in Tennessee at Nuthatch sites", sep=""))
  print(p)
  #fig_name <- paste("data/ebird/", covar, "_covar_tenn.png", sep="")
  #ggsave(fig_name, plot=p, dpi=300, width=15, height=8, units="cm", bg="white", device="png", type="cairo")
}


```
Everything looks good which means we can proceed to modelling. First we have to generate Presence-Absence data. To do so, we are going to nee
to make some assumptions about the parameters.

We will draw values $Y$ from the prior predictive distirbution $p(Y|D, \theta)$. To generate from the ppd, we can do a variety of things. We can either fit a SO model
to the data and use these estimates to generate the PA data. Or we can fit the Presence-only model, and sample from the posterior of this, for each repetition of the PA data. 

```{r data-setup}
# Prepare data for feeding into stan models
link <- "cloglog"
data_reps <- 20
covar_df <- st_drop_geometry(subgrid_covars)
colnames(covar_df)
intensity_covars <- covar_df[c("evi", "elev", "frac_115", "frac_124")]
bias_covars <- covar_df[c("frac_50")]#"duration_minutes", "number_observers",  "effort_distance_km", "time_observations_started", 
k_param_intn <- length(colnames(intensity_covars)) 
k_param_bias <- length(colnames(bias_covars)) 
dim(as.matrix(covar_df))
p_detection <- 0.2
m <- 5
n_surveys <- 5
sites <- nrow(covar_df)
sites
sum(covar_df$counts)
```

```{r fit-ppd}
nuthatch_poisson_process <- '
    data{
      int<lower = 1> N;
      int<lower=0> k_i; // number of predictors for intensity
      int<lower=0> k_b; // number of predictors for bias
      matrix[N, k_i] X; // predictor matrix for intensity
      matrix[N, k_b] Z; // predictor matrix for bias
      array[N] int y;
    }
    parameters{
      real alpha;
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
    }
    model{
      //priors
      target += normal_lpdf(alpha | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(delta | 0,10);

      // likelihood
      // log parameterization so we dont have to exponentiate coefs before passing into poisson distribution
      // https://mc-stan.org/docs/functions-reference/poisson-distribution-log-parameterization.html
      target += poisson_log_lpmf(y | alpha + X * beta + gamma + Z * delta);
    }
    generated quantities{
      //vector[N] lambda_rep;
      //vector[N] lambda_rep_log;
      //vector[N] b_rep;
      //vector[N] lambda_bias;
      array[N] int y_rep;
      //lambda_rep = exp(alpha + beta * X);
      //lambda_rep_log = alpha + beta * X;
      ////.*abs(alpha + beta * X);
      //b_rep = exp(gamma + delta * Z);
      ////.*abs(gamma + delta * Z);
      //// Thin the process
      ////lambda_bias = lambda_rep .* b_rep;
      //lambda_bias = exp(alpha + beta * X + gamma + delta * Z);
      //// User guide has some examples of log poisson generation and poisson PPCs
      y_rep = poisson_log_rng(alpha + X * beta + gamma + Z * delta);
    }
'


model_path <- "nuthatch_poisson_process.stan"
model_string <- nuthatch_poisson_process
write(model_string, model_path)
data_site_occ = list(N=sites, X=intensity_covars, y=covar_df$counts, Z=bias_covars, k_i=k_param_intn, k_b=k_param_bias)
model <- cmdstan_model(model_path) 
fit <- model$sample(data=data_site_occ, seed=13, chains=1, iter_sampling=1000, iter_warmup=100) 
```
Take a look at the fit.
```{r pp-fit}
fit$summary()
all_params <- c("alpha", "gamma", "beta[1]", "beta[2]", "beta[3]", "beta[4]", "delta[1]")
intensity_params <- c("alpha", "beta[1]", "beta[2]", "beta[3]", "beta[4]")
params_intercept <- c("alpha", "gamma")
posterior <- fit$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[1]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[2]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[3]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[4]")) + plot_title

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.9, point_est="mean", pars = c("delta[1]")) + plot_title
print(p_post)

mcmc_intervals(fit$draws(), pars=all_params)
mcmc_hist(fit$draws(), pars = all_params)
mcmc_pairs(fit$draws(), pars=all_params)
mcmc_scatter(fit$draws(), pars=c('alpha', 'gamma'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[2]'))
mcmc_scatter(fit$draws(), pars=c('beta[2]', 'beta[3]'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[4]'))

# Examine the ppd for y
generated_yrep <- fit$draws("y_rep", format="matrix")
dim(generated_yrep)
yrep_means <- colMeans(generated_yrep)
# Create new sf for the preds
subgrid_geo <- st_geometry(subgrid_covars) %>% st_sf()
subgrid_geo$preds <- yrep_means
# Then plot the predictions
pp_pred_rast <- terra::rasterize(vect(subgrid_geo), crop_evi_rast, field="preds")
p <- ggplot() + 
    geom_spatraster(data=pp_pred_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle("Predicted counts at nuthatch sites")
print(p)
```

```{r PA-data}
# Extract the posterior draws from the stan fit
pp_posterior <- fit$draws(intensity_params, format="matrix")
colnames(pp_posterior)
dim(pp_posterior)

generate_ebird_pa <- function(data_reps, surface_data, p_0, pp_posterior, n, sites, link){
  # Generate some random covariate data that will be used to model theta
  # Add a column of 1s to the auxiliary data
  a_intercept <- rep(1, nrow(surface_data))
  surface_data <- cbind(a_intercept, surface_data)
  occupancy_maps <- matrix(nrow=data_reps, ncol=sites)
  Y_detection <- matrix(nrow=data_reps, ncol=sites)
  theta_reps <- matrix(nrow=data_reps, ncol=sites)
  for (r in 1:data_reps){
    # Equivalent way to induce correlation
    #R <- t(chol(corr_matrix)) 
    #theta <- b_0 + sampling_surface$aux*b_1 + R %*% rnorm(sites)
    param_vec <- pp_posterior[r,]
    if (link=="cloglog"){
      g_theta <- 1 - exp(-exp(as.matrix(surface_data)%*%t(param_vec)))
    }
    stopifnot(sites==length(g_theta))
    site_presence <- rbinom(sites, 1, g_theta)
    theta_reps[r, ] <- g_theta
    occupancy_maps[r, ] <- site_presence
    Y_detection[r, ] <- rbinom(sites, n, p_0*site_presence)
  } 
  return(list(occupancy=occupancy_maps, Y=Y_detection, theta=theta_reps))
}

# Generate the data by drawing over the posterior of the parameters
pa_gen <- generate_ebird_pa(data_reps, intensity_covars, p_detection, pp_posterior, n_surveys, sites, link)

# Create plots of the occupancy and probability maps
for(d_i in 1:20){
  occ_map <- st_geometry(subgrid_covars) %>% st_sf()
  occ_map$occ <- pa_gen$occupancy[d_i,]
  occ_map$prob <- pa_gen$theta[d_i,]
  occ_map$counts <- pa_gen$Y[d_i,]
  
  # Then plot the predictions
  occ_rast <- terra::rasterize(vect(occ_map), crop_evi_rast, field="occ")
  p <- ggplot() + 
      geom_spatraster(data=occ_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Occupancy maps generated over Prior Predictive Distribtion")
  print(p)
  
  prob_rast <- terra::rasterize(vect(occ_map), crop_evi_rast, field="prob")
  p <- ggplot() + 
      geom_spatraster(data=prob_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Probability maps generated over Prior Predictive Distribtion")
  print(p)
  
  count_rast <- terra::rasterize(vect(occ_map), crop_evi_rast, field="counts")
  p <- ggplot() + 
      geom_spatraster(data=count_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Counts at sites generated over Prior Predictive Distribtion")
  print(p)
}
```

Check out some other resources about occupancy modelling in stan: UBMS, Richard erickson "occupancy modelling in stan"
```{r stan-model-so}
nuthatch_poisson_process_site_occupancy <- '
    data{
      int<lower = 1> n_pa_sites;
      int<lower = 1> n_po_sites;
      int<lower=0> k_i;
      int<lower=0> k_b;
      int<lower=0, upper=1> model_diag;
      array[n_pa_sites] int Y;
      array[n_pa_sites] int n_surveys;
      matrix[n_pa_sites, k_i] X;
      array[n_po_sites] int PO;
      matrix[n_po_sites, k_i] X_po; // predictor matrix for intensity
      matrix[n_po_sites, k_b] Z_po; // predictor matrix for bias
    }
    parameters{
      real alpha;           // intercepts
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
      real<lower = 0, upper = 1> p;
    }
    model{
      // priors
      target += normal_lpdf(alpha |   0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(delta | 0,10);
      vector[n_pa_sites] g_theta;
      target += poisson_log_lpmf(PO | alpha + X_po*beta + gamma + Z_po*delta);
      g_theta = 1 - exp(-exp(alpha + X*beta));
      for (i in 1:n_pa_sites) {
        if (Y[i] > 0){
          target += log(g_theta[i]*choose(n_surveys[i], Y[i])*(p^Y[i])*(1-p)^(n_surveys[i]-Y[i]));
        }
        else{
          // Compute mixture of no detection and no occupancy
          target += log(g_theta[i]*(1-p)^(n_surveys[i]) + (1 - g_theta[i]));
        }
      }
    }
    generated quantities{
      // Required generated quantities
      // We can generate over all sites, instead of just those being used for PA
      vector[n_po_sites] g_theta_gen;
      array[n_po_sites] int occ_gen; 
      // Note that we do not include gamma here because the 
      // probability of occupancy is just P(N(C_i)>0) = G(alpha + beta X)
      g_theta_gen = 1 - exp(-exp(alpha + X_po*beta));
      // Posterior predictive distribution for occupancy
      occ_gen = bernoulli_rng(g_theta_gen);
    }
'
```
Using the PO data and generated PA data, we can now fit our site-occupancy model. We randomly sample $m$ sites for the PA likelihood.
```{r model-fit}
# We will use the same data as we used to fit the PO model, only now we will also randomly sample a select set of 
# sites to use as PA data in the SO model. Otherwise the set-up stays the same.
site_idx <- sample(which(pa_gen$Y[10,]>0), m, replace=F)
#site_idx <- sample(1:sites, m, replace=F)
site_idx
covars_sampled_sites <- intensity_covars[site_idx, ]
dim(covars_sampled_sites)
covars_sampled_sites
# The PA counts are those generated from the PP parameters above.
selected_counts <- pa_gen$Y[10,site_idx]
selected_counts
n_surveys <- rep(n_surveys, m)
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=covars_sampled_sites, Y=selected_counts, PO=covar_df$counts,
                         X_po=intensity_covars, Z_po=bias_covars, k_i=k_param_intn, k_b=k_param_bias, model_diag=0)
model_path <- "nuthatch_site_occ.stan"
model_string <- nuthatch_poisson_process_site_occupancy
write(model_string, model_path)
model <- cmdstan_model(model_path) 
fit <- model$sample(data=data_site_occ, seed=13, chains=1, iter_sampling=2000, iter_warmup=500) 
```

```{r model-inspect}
#fit$summary()
all_params <- c("alpha", "gamma", "beta[1]", "beta[2]", "beta[3]", "beta[4]", "delta[1]")
intensity_params <- c("alpha", "beta[1]", "beta[2]", "beta[3]", "beta[4]")
params_intercept <- c("alpha", "gamma")
posterior <- fit$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[1]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[2]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[3]")) + plot_title
mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c("beta[4]")) + plot_title

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.9, point_est="mean", pars = c("delta[1]")) + plot_title
print(p_post)

mcmc_intervals(fit$draws(), pars=all_params)
mcmc_hist(fit$draws(), pars = all_params)
mcmc_pairs(fit$draws(), pars=all_params)
mcmc_scatter(fit$draws(), pars=c('alpha', 'gamma'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[2]'))
mcmc_scatter(fit$draws(), pars=c('beta[2]', 'beta[3]'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[4]'))

# Examine the ppd for y
generated_yrep <- fit$draws("g_theta_gen", format="matrix")
dim(generated_yrep)
yrep_means <- colMeans(generated_yrep)
# Create new sf for the preds
subgrid_geo <- st_geometry(subgrid_covars) %>% st_sf()
subgrid_geo$preds <- yrep_means
# Then plot the predictions
occ_pred_rast <- terra::rasterize(vect(subgrid_geo), crop_evi_rast, field="preds")
p <- ggplot() + 
    geom_spatraster(data=occ_pred_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle("Predicted counts at nuthatch sites")
print(p)
```
