---
title: "Site occupancy modelling of Brown-headed Nuthatch Ebird data"
data: today
author: Max Savery
format: pdf
---

```{r set-up}
# Load the ebird data in the way it will be used in the sampling procedure, and fit it to a Bayesian model.
library(sf)
library(terra)
library(tidyterra)
library(sf)
library(auk)
library(tidyverse)
library(lubridate)
library(exactextractr)
library(cmdstanr)
library(bayesplot)


select <- dplyr::select

# Set important global variables 
base_data_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\data\\"
working_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\code\\optimal_design_site_occ"
data_save_dir <- "data/ebird"
save_dir <- file.path(working_dir, data_save_dir)
ebd_download_dir <- "ebd_US_bnhnut_201901_201912_smp_relJul-2023"
lc_path <- "copernicus_landcover/W100N40_PROBAV_LC100_global_v3.0.1_2019-nrt_Discrete-Classification-map_EPSG-4326.tif"
modis_path <- "modis_landcover_dynamics/MCD12Q2.061_EVI_Area_0_doy2019001_aid0001.tif"
elev_path <- "elevation_aster/ASTGTM_NC.003_ASTER_GDEM_DEM_doy2000061_aid0001.tif"

# Source modules
source(file.path(working_dir, "data_utils", "load_ebird_data.R"))
source(file.path(working_dir, "experimental_design_functions.R"))
stan_models_path <- file.path(working_dir, "stan_models", "stan_nuthatch_models.R")
source(stan_models_path)

pp_fit <- FALSE
```

```{r load-initial-data}
data_reps <- 10
data_pack <- read_rds(file.path(save_dir, "data_pack.RDS"))
# Unpack
bias_covars <- data_pack$bias_covars
intensity_covars <- data_pack$intensity_covars
k_param_intn <- data_pack$k_param_intn
k_param_bias <- data_pack$k_param_bias
site_centroids <- data_pack$site_centroids
subgrid <- data_pack$subgrid
state_po_prj <- data_pack$state_po_prj
rast_surface <- rast(rast_surface_path) 
sites <- nrow(intensity_covars)
print(paste("Total sites: ", sites, sep=""))
# Generate po data to match format of simulation script. 
Y_po <- matrix(rep(data_pack$counts, data_reps), nrow=data_reps, ncol=sites, byrow=T)
r_po_data <- list(Y=Y_po)
```

```{r process-covar}
# And just for later use, make a plot with a mark on a specific site
p <- ggplot() + 
  geom_spatraster(data=rast_surface) +
  geom_sf(data = subgrid_covars[c(500, 600, 700),], color=alpha("red",1), fill="orange") + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7, na.value="white") +
  theme_minimal()+
  ggtitle("Labeled nuthatch site")
print(p)

# We can also plot the centroids
bigger_size <- 5/(length(5)) + 1.5
smaller_size <- 1/(length(5)) + 1
site_centroids <- st_centroid(st_geometry(subgrid_covars))
p <- ggplot() + 
  geom_spatraster(data=rast_surface) +
  geom_sf(data = site_centroids[c(1,2,3)], color=alpha("white",1), size=bigger_size) + 
  geom_sf(data = site_centroids[c(100, 200, 300)], color=alpha("hotpink",1), size=smaller_size) + 
  geom_sf(data = site_centroids[400], color=alpha("black",1), size=0.5) + 
  scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.7, na.value="white") +
  theme_minimal()+
  ggtitle("Labeled nuthatch site")
print(p)
```
Everything looks good which means we can proceed to modelling. First we have to generate Presence-Absence data. To do so, we are going to nee
to make some assumptions about the parameters.

We will draw values $Y$ from the prior predictive distirbution $p(Y|D, \theta)$. To generate from the ppd, we can do a variety of things. We can either fit a SO model
to the data and use these estimates to generate the PA data. Or we can fit the Presence-only model, and sample from the posterior of this, for each repetition of the PA data. 

```{r fit-ppd}
if (pp_fit == TRUE){
  # We fit the PP model to the PO data and use the expectation of this posterior to generate the PA data
  fit_point_process_ebird(stan_path, exp_dir, sites, area_a, data_reps, intensity_covars, bias_covars, r_po_data, k_param_intn, k_param_bias, pp_diagnostic=T,
                        subgrid, rast_surface)
}
pp_posterior <- read_rds(file.path(save_dir, "pp_posterior_ebird.RDS"))
print("Intensity fit from point process")
pp_posterior <- matrix(rep(colMeans(pp_posterior), data_reps), nrow=data_reps, byrow = T)
print(pp_posterior)
# Get 1000 samples from the posterior if we have access to the matrix of the MCMC draws from the posterior. 
# Currently the function provides the data rep repeated means of the posterior
#p_index <- sample(c(1:1000), size = 1000, replace = T)
#pp_posterior <- pp_posterior[p_index, ]
#dim(pp_posterior)
```
The above posterior will use the expectation for the parameters instead of sampling from the posterior.

Then generate the PA data 
```{r PA-data}
# Generate the data by drawing over the posterior of the parameters
pa_gen <- generate_ebird_pa(data_reps, area_a, intensity_covars, p_detection, pp_posterior, n_surveys, sites, link)
# Create plots of the occupancy and probability maps
for(d_i in 1:20){
  occ_map <- st_geometry(subgrid) %>% st_sf()
  occ_map$occ <- pa_gen$occupancy[d_i,]
  occ_map$prob <- pa_gen$theta[d_i,]
  occ_map$counts <- pa_gen$Y[d_i,]
  
  # Then plot the predictions
  occ_rast <- terra::rasterize(vect(occ_map), rast_surface, field="occ")
  p <- ggplot() + 
      geom_spatraster(data=occ_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Occupancy maps generated over Prior Predictive Distribtion")
  print(p)
  
  prob_rast <- terra::rasterize(vect(occ_map), rast_surface, field="prob")
  p <- ggplot() + 
      geom_spatraster(data=prob_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Probability maps generated over Prior Predictive Distribtion")
  print(p)
  
  count_rast <- terra::rasterize(vect(occ_map), rast_surface, field="counts")
  p <- ggplot() + 
      geom_spatraster(data=count_rast) +
      scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
      #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
      theme_minimal()+
      ggtitle("Counts at sites generated over Prior Predictive Distribtion")
  print(p)
}
```

Check out some other resources about occupancy modelling in stan: UBMS, Richard erickson "occupancy modelling in stan"

Using the PO data and generated PA data, we can now fit our site-occupancy model. We randomly sample $m$ sites for the PA likelihood.
```{r model-fit}
# We will use the same data as we used to fit the PO model, only now we will also randomly sample a select set of 
# sites to use as PA data in the SO model. Otherwise the set-up stays the same.
#site_idx <- sample(which(pa_gen$Y[10,]>0), m, replace=F)
site_idx <- sample(1:sites, m, replace=F)
site_idx
covars_sampled_sites <- intensity_covars[site_idx, ]
dim(covars_sampled_sites)
covars_sampled_sites
# The PA counts are those generated from the PP parameters above.
# For the sake of illustration we pick 1 dataset
selected_counts <- pa_gen$Y[10,site_idx]
selected_counts
n_surveys <- rep(n_surveys, m)
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, area_a=area_a, X=covars_sampled_sites, Y=selected_counts, PO=covar_df$counts,
                         X_po=intensity_covars, Z_po=bias_covars, k_i=k_param_intn, k_b=k_param_bias, model_diag=0)
model_path <- "nuthatch_site_occ.stan"
model_string <- nuthatch_poisson_process_site_occupancy
write(model_string, model_path)
model <- cmdstan_model(model_path) 
fit <- model$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500) 
```

```{r model-inspect}
#fit$summary()
posterior <- fit$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit$draws(), pars=all_params)
mcmc_hist(fit$draws(), pars = all_params)
mcmc_pairs(fit$draws(), pars=all_params)
mcmc_scatter(fit$draws(), pars=c('alpha', 'gamma'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[2]'))
mcmc_scatter(fit$draws(), pars=c('beta[2]', 'beta[3]'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[4]'))
mcmc_scatter(fit$draws(), pars=c('beta[1]', 'beta[5]'))

# Examine the ppd for y
generated_occ <- fit$draws("occ_gen", format="matrix")
generated_prob <- fit$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)
# Create new sf for the preds
subgrid_geo <- st_geometry(subgrid_covars) %>% st_sf()
subgrid_geo$preds <- occ_means
subgrid_geo$occ_mode <- occ_mode
subgrid_geo$prob_pred <-  prob_means
# Then plot the predictions
occ_pred_rast <- terra::rasterize(vect(subgrid_geo), crop_evi_rast, field="preds")
p <- ggplot() + 
    geom_spatraster(data=occ_pred_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle("Predicted occupancy at nuthatch sites")
print(p)
occ_mode_rast <- terra::rasterize(vect(subgrid_geo), crop_evi_rast, field="occ_mode")
p <- ggplot() + 
    geom_spatraster(data=occ_mode_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle("Predicted occupancy mode at nuthatch sites")
print(p)
# Oops if we have probability less than 0.5 mode is always 0.
prob_pred_rast <- terra::rasterize(vect(subgrid_geo), crop_evi_rast, field="prob_pred")
p <- ggplot() + 
    geom_spatraster(data=prob_pred_rast) +
    scale_fill_viridis_c(begin=0.2, end=1, option="viridis",alpha=0.8, na.value="white") +
    #geom_sf(data = prj_state, color=alpha("black",0.4), fill='transparent', linewidth=0.7) + 
    theme_minimal()+
    ggtitle("Generated occupancy probability at nuthatch sites")
print(p)
```
That summarizes the work needed to generate PA data using the prior predictive distribution from a PP when we have real data.
