---
title: "Simulation of Presence-Absence data when integrating over priors for parameters"
data: today
author: Max Savery
format: pdf
---
  
```{r init-set-up, include=FALSE}
library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(viridis)

working_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\code\\optimal_design_site_occ"
source(file.path(working_dir, "experimental_design_functions.R"))
source(file.path(working_dir, "presence_only_functions.R"))
```
This report is intended to show the deteriorating quality of data generation as we draw from wide priors. We need well-generated PA datasets, but if we draw from the priors of parameters that are too wide, we end up with PA data is overly noisy and will not lead to us being able to easily discriminate models, regardless of the improvements of one model versus another. This is related to the report generating PA data using both an oracle distribution and priors taken from a PP model.

We follow the same procedure as that report, but focus more on the data generation than the model fitting. The two procedures we compare are these:

1. Hold the parameters as fixed or "known" and generate R datasets accordingly. This allows randomness in $y$ but not in $\theta$. Check the variability of this generated data. Then fit an SO 
model on this generated data, and check the Posterior Predictive Distribution (PPD) of the of data per site.    

2. Draw from the priors for each parameter. Generate R datasets accordingly. Check the variability in the datasets. Then fit the SO model, and check the variability in the PPD. We expect the PPD to be very wide, and thus generate noisy estimates for occupancy at any given site.

For each of these procedures, we can compare the oracle model (decoupling PO generation and PA generation) and the sequential data generation model (PA is generated using estimates from PO data)

We can now conduct these experiments.
```{r stan-models}
poisson_process <- '
    data{
      int<lower = 1> N;
      int<lower=0> k_i; // number of predictors for intensity
      int<lower=0> k_b; // number of predictors for bias
      matrix[N, k_i] X; // predictor matrix for intensity
      matrix[N, k_b] Z; // predictor matrix for bias
      array[N] int y;
    }
    parameters{
      real alpha;
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
    }
    model{
      //priors
      target += normal_lpdf(alpha | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(delta | 0,10);

      // likelihood
      // log parameterization so we dont have to exponentiate coefs before passing into poisson distribution
      // https://mc-stan.org/docs/functions-reference/poisson-distribution-log-parameterization.html
      target += poisson_log_lpmf(y | alpha + X * beta + gamma + Z * delta);
    }
    generated quantities{
      array[N] int y_rep;
      y_rep = poisson_log_rng(alpha + X * beta + gamma + Z * delta);
    }
'

poisson_process_site_occupancy <- '
    data{
      int<lower = 1> n_pa_sites;
      int<lower = 1> n_po_sites;
      int<lower=0> k_i;
      int<lower=0> k_b;
      int<lower=0, upper=1> model_diag;
      array[n_pa_sites] int Y;
      array[n_pa_sites] int n_surveys;
      matrix[n_pa_sites, k_i] X;
      array[n_po_sites] int PO;
      matrix[n_po_sites, k_i] X_po; // predictor matrix for intensity
      matrix[n_po_sites, k_b] Z_po; // predictor matrix for bias
    }
    parameters{
      real alpha;           // intercepts
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
      real<lower = 0, upper = 1> p;
    }
    model{
      // priors
      target += normal_lpdf(alpha |   0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(delta | 0,10);
      vector[n_pa_sites] g_theta;
      target += poisson_log_lpmf(PO | alpha + X_po*beta + gamma + Z_po*delta);
      g_theta = 1 - exp(-exp(alpha + X*beta));
      for (i in 1:n_pa_sites) {
        if (Y[i] > 0){
          target += log(g_theta[i]*choose(n_surveys[i], Y[i])*(p^Y[i])*(1-p)^(n_surveys[i]-Y[i]));
        }
        else{
          // Compute mixture of no detection and no occupancy
          target += log(g_theta[i]*(1-p)^(n_surveys[i]) + (1 - g_theta[i]));
        }
      }
    }
    generated quantities{
      // Required generated quantities
      // We can generate over all sites, instead of just those being used for PA
      vector[n_po_sites] g_theta_gen;
      array[n_po_sites] int occ_gen; 
      // Note that we do not include gamma here because the 
      // probability of occupancy is just P(N(C_i)>0) = G(alpha + beta X)
      g_theta_gen = 1 - exp(-exp(alpha + X_po*beta));
      // Posterior predictive distribution for occupancy
      occ_gen = bernoulli_rng(g_theta_gen);
    }
'
```
First set hyperparameters used throughout the data generation and model fitting:
```{r set-vars}
n <- 5
k <- 20
sites <- k^2
area_D <- 400
centroid <- c(10,4)
centroid_2 <- c(20,4)
exp_args <- list(bias_func="exponential")
link_func <- "cloglog"
data_reps <- 10
p_detection <- 0.2
n_chains <- 3
mcmc_iter <- 2000
```
Then generate the simulated sampling surface data
```{r surface-data}
get_bias_surface_misspecified <- function(sampling_grid, centroid, strength){
  x <- seq(1:k)
  y <- seq(1:k)
  # Generate all possible coordinate points
  grid_points <- expand.grid(x, y)
  # Then compute distance from centroid to every other location
  r <- apply(grid_points, 1, function(x, center_coord){sqrt((center_coord[1] - x[1])^2 + (center_coord[2] - x[2])^2)}, center_coord=centroid)
  # Then create z covariate
  x <- exp(-strength*((r)/5))
  x_1 <- qnorm(0.98*x + .01)
  sampling_grid$aux_e <- x_1
  return(sampling_grid)
}

sampling_surface <- get_sampling_surface_donut(k)
sampling_surface <- get_bias_surface_exponential(sampling_surface, centroid)
# Strength of 0.5
sampling_surface <- get_bias_surface_misspecified(sampling_surface, centroid_2, 0.5)
# But for these experiments use 0 misspecification
sampling_surface$aux_e = rep(0, nrow(sampling_surface))

p <- ggplot(sampling_surface, aes(x, y, fill=aux_x)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="X") +
  ggtitle("Initial sampling surface, X covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_z)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Z") +
  ggtitle("Initial sampling surface, Z covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_e)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="E") +
  ggtitle("Initial sampling surface, E covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

```
Define functions for data generation. The PA functions a set of draws or vector of prior values
```{r data-gen-func}
# This function is a little different than in the optimal design code as I am using multiple prior draws
# of params instead of fixed values
generate_pa <- function(data_reps, surface_data, p_0, prior, n, sites, link){
  # Add a column of 1s to the auxiliary data
  a_intercept <- rep(1, nrow(surface_data))
  surface_data <- cbind(a_intercept, surface_data)
  occupancy_maps <- matrix(nrow=data_reps, ncol=sites)
  Y_detection <- matrix(nrow=data_reps, ncol=sites)
  theta_reps <- matrix(nrow=data_reps, ncol=sites)
  for (r in 1:data_reps){
    param_vec <- as.matrix(prior[r,])
    if (link=="cloglog"){
      g_theta <- 1 - exp(-exp(as.matrix(surface_data)%*%matrix(param_vec)))
    }else{
      stop("Link not available")
    }
    stopifnot(sites==length(g_theta))
    site_presence <- rbinom(sites, 1, g_theta)
    theta_reps[r, ] <- g_theta
    occupancy_maps[r, ] <- site_presence
    Y_detection[r, ] <- rbinom(sites, n, p_0*site_presence)
  } 
  return(list(occupancy=occupancy_maps, Y=Y_detection, theta=theta_reps))
}

generate_ppp_data_misspecified <- function(surface_data, params, n_sites, data_reps, corr_matrix, gp_bool, area_D){
  # Generate some random covariate data per site on a grid that will be used to model lambda and b
  # There will be r (data_reps) datasets replicated, for use with the exchange 
  # algorithm/monte carlo integration and to maintain efficiency in the parallel 
  # function computation of V
  lambda <- (area_D / n_sites) * exp(params$alpha + params$beta*surface_data$aux_x)
  eta <- exp(params$gamma + params$delta*surface_data$aux_z + params$epsilon*surface_data$aux_e)
  b <- eta / (1 + eta)
  if (gp_bool==T){
    gp <- exp(rnorm(n_sites, 0, corr_matrix))
    Y_counts <- rpois(n_sites, lambda * b * gp)
  }
  else{
    Y_counts <- rpois(n_sites, lambda * b)
  }
  # Generate coordinate points within each site for each observation.
  Y_coords <- generate_coords(surface_data, Y_counts)
  # Create matrix of data reps, though for po data we just copy everything,
  # as opposed to random generation for each rep.
  Y_po <- matrix(rep(Y_counts, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  lambdas <- matrix(rep(lambda, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  biases <- matrix(rep(b, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  return(list(Y=Y_po, lambda=lambdas, bias=biases, Y_coords=Y_coords))
}

```
Having set up our sampling space and data generating processes, we can draw our parameters. Let's first use fixed values or expectations as our known parameter values, with the oracle model.
```{r oracle-data-gen}
alpha <- -2
beta <- 0.5
alpha_draws <- rep(alpha, data_reps)
beta_draws <- rep(beta, data_reps)
pa_param_draws <- cbind(alpha_draws, beta_draws)
# Use mean as fixed values for PO params 
gamma <- 1
delta <- 2
epsilon <- 1.5
po_param_vec <- list(alpha=alpha, beta=beta, gamma=gamma, delta=delta, epsilon=1.5)

# Using pre-defined funcky for PO
po_gen <- generate_ppp_data_misspecified(sampling_surface, po_param_vec, sites, data_reps, corr_matrix=NA, gp_bool=FALSE, area_D)
# And the function defined above for the PA data
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pa_param_draws, n, sites, link_func)

# Check the generated_data
text_size <- 10
# Plot PA data
for (d in 1:data_reps){
  subtitle <- paste("alpha=", alpha_draws[d],", beta=", beta_draws[d], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}

# Then the PO data. Only 1 set of repeated po data. The PO data is all the same over the rows (repeated for efficiency in OD)
p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$lambda[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Lambda") +
  ggtitle("Generated intensity per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$bias[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, "Bias") +
  ggtitle("Generated bias per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

thinned_intensity <- po_gen$lambda[data_reps,]*po_gen$bias[data_reps,]
p <- ggplot() +
  geom_tile(sampling_surface, mapping=aes(x, y, fill=thinned_intensity, width=1, height=1), alpha=.6) + 
  scale_fill_viridis(discrete=FALSE, name="L*b") +
  ggtitle("Generated PP thinning per site, \nincluding observed PO individuals") +
  geom_point(data=po_gen$Y_coords, mapping=aes(x=x, y=y), size=2, col="white") +
  theme(panel.grid.minor = element_line(colour="white")) +
  scale_y_continuous(breaks = seq(0, 20, 1)) +
  scale_x_continuous(breaks = seq(0, 20, 1)) +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)
```
First compile the model for both experiments
```{r stan-compile}
model_path <- "pp_so_for_sim.stan"
model_string <- poisson_process_site_occupancy
write(model_string, model_path)
model_so <- cmdstan_model(model_path) 
```

One question is what effect does sampling $m$ subset of sites versus using the full set of sites in the the SO model in the context of this experiment.

In the model below, it is fit to to just the first replication of PA data here. This means that we are not taking the 
uncertainty of Y into account in these experiments. We just consider a fixed dataset for fixed parameters. 
```{r oracle-stan-fit}
# Using all sites for now
# We could also subsample as in OD
m <- sites
#site_idx <- sample(which(pa_gen$Y[10,]>0), m, replace=F)
#site_idx <- sample(1:sites, m, replace=F)
#site_idx
covars_sampled_sites <- sampling_surface$aux_x
# The PA counts are those generated from the PP parameters above.
# For the sake of illustration we pick 1 dataset
#selected_counts <- pa_gen$Y[10,site_idx]
#selected_counts
# n is for number of visits
n_surveys <- rep(n, sites)
k_i <- 1
k_b <- 1

data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,], X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
# Using the earlier compiled model, fit to PO+PA data
fit_so_1_fixed <- model_so$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
```
```{r oracle-diag}
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_1_fixed$summary(all_params)
posterior <- fit_so_1_fixed$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

# Examine the ppd for y
generated_occ <- fit_so_1_fixed$draws("occ_gen", format="matrix")
generated_prob <- fit_so_1_fixed$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```

In the next experiment, we use the same generated PO dataset, using the same values of the parameters. We then fit the PP model to this dataset, and use the mean estimates from this model to generate the PA data.
```{r po-est}
data_site_occ = list(N=sites, X=as.matrix(sampling_surface$aux_x), y=po_gen$Y[1,], Z=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b)
model_path <- "poisson_process_sim.stan"
model_string <- poisson_process
write(model_string, model_path)
model_pp <- cmdstan_model(model_path) 
fit_pp_fixed <- model_pp$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
```
Check the diagnostic fit. We should recover the parameters pretty well here, without any misspecification, though there is the correlation between alpha and gamma. Compare these estimates to the true means.
```{r pp-diag}
fit_pp_fixed$summary(all_params)
posterior <- fit_pp_fixed$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_pp_fixed$draws(), pars=all_params)
mcmc_pairs(fit_pp_fixed$draws(), pars=all_params)
mcmc_scatter(fit_pp_fixed$draws(), pars=c('alpha', 'gamma'))

generated_y <- fit_pp_fixed$draws("y_rep", format="matrix")
yrep_mean <- colMeans(generated_y)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, yrep_mean=yrep_mean)

p <- ggplot(ppd_df, aes(x, y, fill=yrep_mean)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD Counts per site")
print(p)
```

```{r pa-gen-with-pp}
pp_posterior_draws <- fit_pp_fixed$draws(c("alpha", "beta[1]"), format="matrix")
# Just use the means from the posterior
pp_posterior <- matrix(rep(colMeans(pp_posterior_draws), data_reps), nrow=data_reps, byrow = T)
stopifnot(dim(pa_param_draws)==dim(pp_posterior))

# Generate a new set of pa data reps
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pp_posterior, n, sites, link_func)

# Check the generated_data over the iterations
text_size <- 10
for (d in 1:data_reps){
  subtitle <- paste("alpha=", pp_posterior[d, 1],", beta=", pp_posterior[d, 2], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}
```
Using the newly generated PA data, we fit another SO model
```{r estimated-po-stan-fit}
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,],
                         X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
# Using the earlier compiled model, fit to PO+PA data
fit_so_2_fixed <- model_so$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500)
```
Check the fit of this sequentially generated model.
```{r estimated-diag}
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_2_fixed$summary(all_params)
posterior <- fit_so_2_fixed$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_so_2_fixed$draws(), pars=all_params)
mcmc_hist(fit_so_2_fixed$draws(), pars = all_params)
mcmc_pairs(fit_so_2_fixed$draws(), pars=all_params)
mcmc_scatter(fit_so_2_fixed$draws(), pars=c('alpha', 'gamma'))

# Examine the ppd for y
generated_occ <- fit_so_2_fixed$draws("occ_gen", format="matrix")
generated_prob <- fit_so_2_fixed$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```
That finishes fitting the models using fixed values (expectations) of the parameters. We then proceed to repeat these experiments but drawing from the priors of the parameters instead.

We actually only draw from the priors for alpha and beta here and use these only to generate the PA data. Let's take a look at that first and then consider if we need to draw from them to generate the PO data.
```{r oracle-stan-fit-prior-draws}
alpha <- -2
beta <- 0.5
# Draw from priors
alpha_draws <- rnorm(data_reps, alpha, 0.5)
beta_draws <- rnorm(data_reps, beta, 0.25)
pa_param_draws <- cbind(alpha_draws, beta_draws)

# Use mean as fixed values for PO params 
gamma <- 0.1
delta <- 2
epsilon <- 1.5
po_param_vec <- list(alpha=alpha, beta=beta, gamma=gamma, delta=delta, epsilon=epsilon)

# Using pre-defined funcky for PO
po_gen <- generate_ppp_data_misspecified(sampling_surface, po_param_vec, sites, data_reps, corr_matrix=NA, gp_bool=FALSE, area_D)
# And the function defined above for the PA data
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pa_param_draws, n, sites, link_func)

# Check the generated_data
text_size <- 10
# Plot PA data
for (d in 1:data_reps){
  subtitle <- paste("alpha=", alpha_draws[d],", beta=", beta_draws[d], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}

# Then the PO data. Only 1 set of repeated po data. The PO data is all the same over the rows (repeated for efficiency in OD)
p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$lambda[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Lambda") +
  ggtitle("Generated intensity per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$bias[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, "Bias") +
  ggtitle("Generated bias per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

thinned_intensity <- po_gen$lambda[data_reps,]*po_gen$bias[data_reps,]
p <- ggplot() +
  geom_tile(sampling_surface, mapping=aes(x, y, fill=thinned_intensity, width=1, height=1), alpha=.6) + 
  scale_fill_viridis(discrete=FALSE, name="L*b") +
  ggtitle("Generated PP thinning per site, \nincluding observed PO individuals") +
  geom_point(data=po_gen$Y_coords, mapping=aes(x=x, y=y), size=2, col="white") +
  theme(panel.grid.minor = element_line(colour="white")) +
  scale_y_continuous(breaks = seq(0, 20, 1)) +
  scale_x_continuous(breaks = seq(0, 20, 1)) +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

m <- sites
covars_sampled_sites <- sampling_surface$aux_x
n_surveys <- rep(n, sites)
k_i <- 1
k_b <- 1
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,], X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
fit_so_1_sampled <- model_so$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_1_sampled$summary(all_params)
posterior <- fit_so_1_sampled$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

# Examine the ppd for y
generated_occ <- fit_so_1_sampled$draws("occ_gen", format="matrix")
generated_prob <- fit_so_1_sampled$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```

Then, we fit the PP -> PA -> SO models but draw from the posterior of the PP model when generating the R PA datasets
```{r po-est-sampled}
data_site_occ = list(N=sites, X=as.matrix(sampling_surface$aux_x), y=po_gen$Y[1,], Z=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b)
model_path <- "poisson_process_sim.stan"
model_string <- poisson_process
write(model_string, model_path)
model_pp <- cmdstan_model(model_path) 
fit_pp_sampled <- model_pp$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
# Get the summary
fit_pp_sampled$summary(all_params)
posterior <- fit_pp_sampled$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_pp_sampled$draws(), pars=all_params)
mcmc_pairs(fit_pp_sampled$draws(), pars=all_params)
mcmc_scatter(fit_pp_sampled$draws(), pars=c('alpha', 'gamma'))

generated_y <- fit_pp_sampled$draws("y_rep", format="matrix")
yrep_mean <- colMeans(generated_y)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, yrep_mean=yrep_mean)

p <- ggplot(ppd_df, aes(x, y, fill=yrep_mean)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD Counts per site")
print(p)
pp_posterior_draws <- fit_pp_sampled$draws(c("alpha", "beta[1]"), format="matrix")
# Sample from the posterior now
p_index <- sample(c(1:(n_chains*mcmc_iter)), size = data_reps, replace = T)
pp_posterior <- pp_posterior_draws[p_index, ]
stopifnot(dim(pa_param_draws)==dim(pp_posterior))
# Generate a new set of pa data reps
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pp_posterior, n, sites, link_func)

# Check the generated_data over the iterations
text_size <- 10
for (d in 1:data_reps){
  subtitle <- paste("alpha=", pp_posterior[d, 1],", beta=", pp_posterior[d, 2], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}
```
Using the newly generated PA data, we fit another SO model
```{r estimated-po-stan-fit-sampled}
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,],
                         X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
# Using the earlier compiled model, fit to PO+PA data
fit_so_2_sampled <- model_so$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500)
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_2_sampled$summary(all_params)
posterior <- fit_so_2_sampled$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_so_2_sampled$draws(), pars=all_params)
mcmc_hist(fit_so_2_sampled$draws(), pars = all_params)
mcmc_pairs(fit_so_2_sampled$draws(), pars=all_params)
mcmc_scatter(fit_so_2_sampled$draws(), pars=c('alpha', 'gamma'))

# Examine the ppd for y
generated_occ <- fit_so_2_sampled$draws("occ_gen", format="matrix")
generated_prob <- fit_so_2_sampled$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```
Let's take a look at the estimates and the PPDs. In the estimates, we expect the oracle models to be more accurate, and the models with fixed parameters to have smaller SD on the parameters. In the PPDs, the fixed models should generate a sharper gaussian intensity donut. 
```{r compare}
print(fit_so_1_fixed$summary(all_params))
print(fit_so_1_sampled$summary(all_params))
print(fit_so_2_fixed$summary(all_params))
print(fit_so_2_sampled$summary(all_params))

# Take a look at all the respective PPDs
generated_occ <- fit_so_1_fixed$draws("occ_gen", format="matrix")
generated_prob <- fit_so_1_fixed$draws("g_theta_gen", format="matrix")
occ_means <- colMeans(generated_occ)
prob_means <- colMeans(generated_prob)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_rep=occ_means, theta_rep=prob_means)
# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site for fixed data-gen SO oracle model")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site for fixed data-gen SO oracle model")
print(p)

generated_occ <- fit_so_1_sampled$draws("occ_gen", format="matrix")
generated_prob <- fit_so_1_sampled$draws("g_theta_gen", format="matrix")
occ_means <- colMeans(generated_occ)
prob_means <- colMeans(generated_prob)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_rep=occ_means, theta_rep=prob_means)
# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site for sampled data-gen SO oracle model")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site for sampled data-gen SO oracle model")
print(p)

generated_occ <- fit_so_2_fixed$draws("occ_gen", format="matrix")
generated_prob <- fit_so_2_fixed$draws("g_theta_gen", format="matrix")
occ_means <- colMeans(generated_occ)
prob_means <- colMeans(generated_prob)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_rep=occ_means, theta_rep=prob_means)
# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site for fixed data-gen SO sequential model")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site for fixed data-gen SO sequential model")
print(p)

generated_occ <- fit_so_2_sampled$draws("occ_gen", format="matrix")
generated_prob <- fit_so_2_sampled$draws("g_theta_gen", format="matrix")
occ_means <- colMeans(generated_occ)
prob_means <- colMeans(generated_prob)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_rep=occ_means, theta_rep=prob_means)
# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site for sampled data-gen SO sequential model")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site for sampled data-gen SO sequential model")
print(p)
```
This shows that when we pull from the prior, indeed the occupancy and occupancy probability estimates will be somewhat diluted than when we know the true values of the parameters. However, this is for only one draw of PA data. Let's take the expectation per site of the PPD for occupancy from the SO models. Here I'll just use the oracle models since they are faster to fit the than two-model PP-SO approach.
```{r expect-over-pa}
data_reps <- 100
n_chains <- 1
alpha <- -2
beta <- 0.5
alpha_fixed <- rep(alpha, data_reps)
beta_fixed <- rep(beta, data_reps)
pa_param_fixed <- cbind(alpha_fixed, beta_fixed)
alpha_draws <- rnorm(data_reps, alpha, 0.5)
beta_draws <- rnorm(data_reps, beta, 0.25)
pa_param_draws <- cbind(alpha_draws, beta_draws)

# Use mean as fixed values for PO params 
gamma <- 0.1
delta <- 2
epsilon <- 1.5
po_param_vec <- list(alpha=alpha, beta=beta, gamma=gamma, delta=delta, epsilon=1.5)
# Using pre-defined funcky for PO
po_gen <- generate_ppp_data_misspecified(sampling_surface, po_param_vec, sites, data_reps, corr_matrix=NA, gp_bool=FALSE, area_D)
# And the function defined above for the PA data
pa_gen_param_fixed <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pa_param_fixed, n, sites, link_func)
pa_gen_param_draws <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pa_param_draws, n, sites, link_func)
# Data holding
occ_matrix_fixed <- matrix(nrow=data_reps, ncol=sites)
occ_matrix_sampled <- matrix(nrow=data_reps, ncol=sites)
for (d_i in 1:data_reps){
  print(paste("On data rep: ", d_i, sep=""))
  # Fixed data
  data_site_occ_fixed = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen_param_fixed$Y[1, ], PO=po_gen$Y[1,], X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
  # Sampled data
  data_site_occ_sampled = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen_param_draws$Y[1, ], PO=po_gen$Y[1,], X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
  # Using the earlier compiled model, fit to PO+PA data
  fit_so_fixed <- model_so$sample(data=data_site_occ_fixed, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500)
  fit_so_sampled <- model_so$sample(data=data_site_occ_sampled, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500)
  # Examine the ppd for y
  generated_occ_fixed <- fit_so_fixed$draws("occ_gen", format="matrix")
  generated_occ_sampled <- fit_so_sampled$draws("occ_gen", format="matrix")
  occ_means_fixed <- colMeans(generated_occ_fixed)
  occ_means_sampled <- colMeans(generated_occ_sampled)
  occ_matrix_fixed[d_i, ] <- occ_means_fixed
  occ_matrix_sampled[d_i, ] <- occ_means_sampled
}
```
Then plot the results by taking the expectation over the data reps.
```{r plot-expectations}
# Plot generated ys and thetas on the sampling grid
occ_fixed_mean <- colMeans(occ_matrix_fixed)
occ_sampled_mean <- colMeans(occ_matrix_sampled)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_fixed=occ_fixed_mean, occ_sampled=occ_sampled_mean)
p <- ggplot(ppd_df, aes(x, y, fill=occ_fixed)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("Expected PPD occupancy per site for fixed data-gen SO oracle model")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_sampled)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("Expected PPD occupancy per site for sampled data-gen SO oracle model")
print(p)
```
The results of these simulations show that the expected occupancy status decreases when we integrate over prior information. On one hand, this ins somewhat obvious, as incorporating more uncertainty from prior information leads us to consider particular events less likely. However, the effect of this depends on the vagueness of the priors. The more poorly estimated our priors our, the worse we can estimate our occupancy, and will be more uncertain about the occupancy at any given site. The effect of this is that we must be careful when choosing our priors, as if they are too vague or badly estimated then we will not get useful occupancy estimates and it will be difficult to distinguish models in an experimental design algorithm where we will rely on the occupancy estimates to minimize a loss function.
