---
title: "Simulation of Integration over Gamma"
data: today
author: Max Savery
format: pdf
---
  
```{r init-set-up, include=FALSE}
library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(viridis)

working_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\code\\optimal_design_site_occ"
source(file.path(working_dir, "experimental_design_functions.R"))
source(file.path(working_dir, "presence_only_functions.R"))
```
This report is intended to show the an approach for fitting the posterior of a Point Process model by integrating over the gamma prior.

This qmd is being converted from with_without prior qmd.

```{r stan-models}
poisson_process <- '
    data{
      int<lower = 1> N;
      int<lower=0> k_i; // number of predictors for intensity
      int<lower=0> k_b; // number of predictors for bias
      matrix[N, k_i] X; // predictor matrix for intensity
      matrix[N, k_b] Z; // predictor matrix for bias
      array[N] int y;
    }
    parameters{
      real alpha;
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
    }
    model{
      //priors
      target += normal_lpdf(alpha | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(delta | 0,10);

      // likelihood
      // log parameterization so we dont have to exponentiate coefs before passing into poisson distribution
      // https://mc-stan.org/docs/functions-reference/poisson-distribution-log-parameterization.html
      target += poisson_log_lpmf(y | alpha + X * beta + gamma + Z * delta);
    }
    generated quantities{
      array[N] int y_rep;
      y_rep = poisson_log_rng(alpha + X * beta + gamma + Z * delta);
    }
'

poisson_process_gamma_constant <- '
    data{
      int<lower = 1> N;
      real gamma;
      int<lower=0> k_i; // number of predictors for intensity
      int<lower=0> k_b; // number of predictors for bias
      matrix[N, k_i] X; // predictor matrix for intensity
      matrix[N, k_b] Z; // predictor matrix for bias
      array[N] int y;
    }
    parameters{
      real alpha;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
    }
    model{
      //priors
      target += normal_lpdf(alpha | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(delta | 0,10);

      // likelihood
      // log parameterization so we dont have to exponentiate coefs before passing into poisson distribution
      // https://mc-stan.org/docs/functions-reference/poisson-distribution-log-parameterization.html
      target += poisson_log_lpmf(y | alpha + X * beta + gamma + Z * delta);
    }
    generated quantities{
      array[N] int y_rep;
      y_rep = poisson_log_rng(alpha + X * beta + gamma + Z * delta);
    }
'
```
First set hyperparameters used throughout the data generation and model fitting:
```{r set-vars}
n <- 5
k <- 20
sites <- k^2
area_D <- 400
centroid <- c(10,4)
exp_args <- list(bias_func="exponential")
link_func <- "cloglog"
data_reps <- 10
p_detection <- 0.2
n_chains <- 3
mcmc_iter <- 2000
```
Then generate the simulated sampling surface data
```{r surface-data}
sampling_surface <- get_sampling_surface_donut(k)
sampling_surface <- get_bias_surface_exponential(sampling_surface, centroid)
# But for these experiments use 0 misspecification
p <- ggplot(sampling_surface, aes(x, y, fill=aux_x)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="X") +
  ggtitle("Initial sampling surface, X covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_z)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Z") +
  ggtitle("Initial sampling surface, Z covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)
```
Define functions for data generation. The PA functions a set of draws or vector of prior values
```{r data-gen-func}
# This function is a little different than in the optimal design code as I am using multiple prior draws
# of params instead of fixed values
generate_pa <- function(data_reps, surface_data, p_0, prior, n, sites, link){
  # Add a column of 1s to the auxiliary data
  a_intercept <- rep(1, nrow(surface_data))
  surface_data <- cbind(a_intercept, surface_data)
  occupancy_maps <- matrix(nrow=data_reps, ncol=sites)
  Y_detection <- matrix(nrow=data_reps, ncol=sites)
  theta_reps <- matrix(nrow=data_reps, ncol=sites)
  for (r in 1:data_reps){
    param_vec <- as.matrix(prior[r,])
    if (link=="cloglog"){
      g_theta <- 1 - exp(-exp(as.matrix(surface_data)%*%matrix(param_vec)))
    }else{
      stop("Link not available")
    }
    stopifnot(sites==length(g_theta))
    site_presence <- rbinom(sites, 1, g_theta)
    theta_reps[r, ] <- g_theta
    occupancy_maps[r, ] <- site_presence
    Y_detection[r, ] <- rbinom(sites, n, p_0*site_presence)
  } 
  return(list(occupancy=occupancy_maps, Y=Y_detection, theta=theta_reps))
}

generate_ppp_data <- function(surface_data, params, n_sites, data_reps, corr_matrix, gp_bool, area_D){
  # Generate some random covariate data per site on a grid that will be used to model lambda and b
  # There will be r (data_reps) datasets replicated, for use with the exchange 
  # algorithm/monte carlo integration and to maintain efficiency in the parallel 
  # function computation of V
  lambda <- (area_D / n_sites) * exp(params$alpha + params$beta*surface_data$aux_x)
  eta <- exp(params$gamma + params$delta*surface_data$aux_z)
  b <- eta / (1 + eta)
  if (gp_bool==T){
    gp <- exp(rnorm(n_sites, 0, corr_matrix))
    Y_counts <- rpois(n_sites, lambda * b * gp)
  }
  else{
    Y_counts <- rpois(n_sites, lambda * b)
  }
  # Generate coordinate points within each site for each observation.
  Y_coords <- generate_coords(surface_data, Y_counts)
  # Create matrix of data reps, though for po data we just copy everything,
  # as opposed to random generation for each rep.
  Y_po <- matrix(rep(Y_counts, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  lambdas <- matrix(rep(lambda, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  biases <- matrix(rep(b, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  return(list(Y=Y_po, lambda=lambdas, bias=biases, Y_coords=Y_coords))
}

```
Having set up our sampling space and data generating processes, we can now compare the effect of integrating over gamma during the MCMC fitting.

We first generate the PO data based on some set params.
```{r gen-po}
alpha <- -2
beta <- 0.5
gamma <- 0.1
delta <- 2
k_i <- 1
k_b <- 1
alpha_draws <- rep(alpha, data_reps)
beta_draws <- rep(beta, data_reps)
pa_param_draws <- cbind(alpha_draws, beta_draws)
# Use mean as fixed values for PO params 
po_param_vec <- list(alpha=alpha, beta=beta, gamma=gamma, delta=delta)

# Using pre-defined funcky for PO
po_gen <- generate_ppp_data(sampling_surface, po_param_vec, sites, data_reps, corr_matrix=NA, gp_bool=FALSE, area_D)

text_size <- 10
# Then the PO data. Only 1 set of repeated po data. The PO data is all the same over the rows (repeated for efficiency in OD)
p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$lambda[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Lambda") +
  ggtitle("Generated intensity per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$bias[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, "Bias") +
  ggtitle("Generated bias per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

thinned_intensity <- po_gen$lambda[data_reps,]*po_gen$bias[data_reps,]
p <- ggplot() +
  geom_tile(sampling_surface, mapping=aes(x, y, fill=thinned_intensity, width=1, height=1), alpha=.6) + 
  scale_fill_viridis(discrete=FALSE, name="L*b") +
  ggtitle("Generated PP thinning per site, \nincluding observed PO individuals") +
  geom_point(data=po_gen$Y_coords, mapping=aes(x=x, y=y), size=2, col="white") +
  theme(panel.grid.minor = element_line(colour="white")) +
  scale_y_continuous(breaks = seq(0, 20, 1)) +
  scale_x_continuous(breaks = seq(0, 20, 1)) +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)
```
Then fit the model with both gamma and alpha
```{r pp-fit-no-integration}
data_site_occ = list(N=sites, X=as.matrix(sampling_surface$aux_x), y=po_gen$Y[1,], Z=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b)
model_path <- "poisson_process_sim.stan"
model_string <- poisson_process
write(model_string, model_path)
model_pp <- cmdstan_model(model_path) 
fit_pp_gamma <- model_pp$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
```
Check the diagnostic fit. 
```{r pp-diag-1}
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
print(fit_pp_gamma$summary(all_params))
posterior <- fit_pp_gamma$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_pp_gamma$draws(), pars=all_params)
mcmc_pairs(fit_pp_gamma$draws(), pars=all_params)
mcmc_scatter(fit_pp_gamma$draws(), pars=c('alpha', 'gamma'))

generated_y <- fit_pp_gamma$draws("y_rep", format="matrix")
yrep_mean <- colMeans(generated_y)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, yrep_mean=yrep_mean)

p <- ggplot(ppd_df, aes(x, y, fill=yrep_mean)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD Counts per site")
print(p)
```

```{r pa-gen-with-pp-1}
pp_posterior_draws <- fit_pp_gamma$draws(c("alpha", "beta[1]"), format="matrix")
# Just use the means from the posterior
pp_posterior <- matrix(rep(colMeans(pp_posterior_draws), data_reps), nrow=data_reps, byrow = T)
print(pp_posterior)
stopifnot(dim(pa_param_draws)==dim(pp_posterior))

# Generate a new set of pa data reps
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pp_posterior, n, sites, link_func)

# Check the generated_data over the iterations
text_size <- 10
for (d in 1:data_reps){
  subtitle <- paste("alpha=", pp_posterior[d, 1],", beta=", pp_posterior[d, 2], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}
```
That is the data generation when we seperately estimate $\alpha$ and $\gamma$. We next observe the effect of integrating over $\gamma$
```{r compile-pp}
model_path <- "poisson_process_sim_gamma.stan"
model_string <- poisson_process_gamma_constant
write(model_string, model_path)
model_pp <- cmdstan_model(model_path) 
```
Now we run the simulation, drawing from a uniform prior for gamma, fixing gamma in the PP model, and fitting the posterior of $\alpha$. We repeat this PP fitting and alpha estimation over all draws from the gamma prior, and then average the posterior mean of $\alpha$ from all of these runs.

```{r pp-fit-integration}
gamma_reps <- 200
gamma_sample <- runif(gamma_reps, -1, 1)
all_params <- c("alpha", "beta[1]", "delta[1]")
# 3 cols for alpha, beta, and delta (gamma is not a parameter)
pp_rep <- matrix(nrow=gamma_reps, ncol=length(all_params))
```
```{r run-sim, message=FALSE, results='hide', echo=FALSE}
for (i in 1:gamma_reps){
  data_site_occ = list(gamma=gamma_sample[i], N=sites, X=as.matrix(sampling_surface$aux_x), y=po_gen$Y[1,], Z=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b)
  fit_pp_gamma_const <- model_pp$sample(data=data_site_occ, seed=13, chains=n_chains, iter_sampling=mcmc_iter, iter_warmup=500) 
  fit_means <- fit_pp_gamma_const$summary(all_params)$mean
  pp_rep[i,] <- fit_means
}
```
```{r sim output}
print("Mean of alpha averaged over gamma")
print(colMeans(pp_rep))
print(pp_rep)
```
Check the diagnostic (final) fit. 
```{r pp-diag-2}
fit_pp_gamma_const$summary(all_params)
posterior <- fit_pp_gamma_const$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = "alpha") + plot_title
print(p_post)

mcmc_intervals(fit_pp_gamma_const$draws(), pars=all_params)
mcmc_pairs(fit_pp_gamma_const$draws(), pars=all_params)

generated_y <- fit_pp_gamma_const$draws("y_rep", format="matrix")
yrep_mean <- colMeans(generated_y)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, yrep_mean=yrep_mean)

p <- ggplot(ppd_df, aes(x, y, fill=yrep_mean)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD Counts per site")
print(p)
```

```{r pa-gen-with-pp-2}
# Just use the means from the posterior
pp_posterior <- matrix(rep(colMeans(pp_rep[,1:2]), data_reps), nrow=data_reps, byrow = T)
print(pp_posterior)
stopifnot(dim(pa_param_draws)==dim(pp_posterior))

# Generate a new set of pa data reps
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pp_posterior, n, sites, link_func)

# Check the generated_data over the iterations
text_size <- 10
for (d in 1:data_reps){
  subtitle <- paste("alpha=", pp_posterior[d, 1],", beta=", pp_posterior[d, 2], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}
```
