---
title: "Simulation of Presence-Absence data when integrating over priors for parameters"
data: today
author: Max Savery
format: pdf
---
  
```{r init-set-up, include=FALSE}
library(tidyverse)
library(cmdstanr)
library(bayesplot)
library(viridis)

working_dir <- "C:\\Users\\msavery\\OneDrive - UGent\\Documents\\ghent_phd_spatial_doe\\code\\optimal_design_site_occ"
source(file.path(working_dir, "experimental_design_functions.R"))
source(file.path(working_dir, "presence_only_functions.R"))
# Bool to draw over priors for all experiments
over_prior <- FALSE
```

This report is intented to show the detoriating quality of data generation as we draw from wide priors. We need well-generated PA datasets, but if we draw from the priors of parameters that are too wide, we end up with PA data is overly noisy and will not lead to us being able to easily discriminate models, regardless of the improvements of one model versus another. This is related to the report generating PA data using both an oracle distribution and priors taken from a PP model.

EDIT:
The theoretical set up takes place in the context of Bayesian Optimal Design. We need to generate data according to different designs, and to generate the data, we need to draw from the predictive distribution. But in the fully Bayesian sense, we don't have the marginal predictive distribution $p(y|d)$ and need to simulate from it. To do so, we can fit a Bayesian model (the posterior will be used as the prior) and draw from the prior predictive distribution in this way. 

In the optimal design set-up, we define the expectation of our utility as

$$
E[U(d)] = \int_y \int_{\theta} u(d, y, \theta)p(\theta|y)p(y|d)d\theta dy
$$
In this Bayesian formulation, we integrate both over $\theta$ and $y$. Since this integral is not tractable in many applications, we need computational methods to approximate the integral. One approach is to compute the Posterior Predictive Distribution (PPD) integrated over $\theta$ by fitting a Bayesian model for $\theta$ and $y$ and drawing $\tilde{y}$ from the PPD. The draw of $\tilde{y}$ can then be used to compute the loss in the utility function. 

To integrate over $y$, we need to repeatedly draw over the marginal distribution $p(y|\theta)$. This can be done by computing the prior predictive distribution of y:
$$
p(y|d) = \int p(y|\theta)p(\theta)d\theta
$$
This requires us to define a prior for $\theta$ and generate y according to it's conditional on $\theta$ (this needs to be known). However, consider if the prior for $\theta$ is based on prior data. In that case we can fit a bayesian model to the prior data, compute $p(\theta)$ as the posterior $p(\theta|y_{prior})\propto p(y_{prior}|\theta_0)p(\theta_0)$, and use draws from this posterior to generate draws from the prior predictive of $y$. We still have to specify an initial prior for $\theta$ $p(\theta_0)$, but such is the sequential nature of Bayesian models. 

Now, we can move into the context of occupancy data. Here, $y_{prior}$ is Presence-Only (PO) data, $y$ is Presence-Absence (PA) data, and $\theta$ refers to a vector of parameters used to model the occupancy probability.

TODO: Discuss using PP as prior and how PA and PO relate

The problem we run into is that we need to be able to draw from $p(y|d)$ before we fit our site-occupancy model. But we do not have a point estimate of $\theta$ or known distribution of $\theta$ to 
draw from. One possibility is to fit a point process model using the PO data to generate a posterior for $\theta$. Then we can draw from this posterior and generate sets of $y$, the PA data. However, in our utility function, we will AGAIN use the PO data as a prior when we integrate over theta. Any biases present in our PO data will effect the generation of the PA data. For example, if our PO data is gathered in convenient locations near urban centers AND this sampling bias is not considered in the estimation of $b$, such bias will show up as underestimating the occupancy of other areas. In the utility function, sites where occupancy has been underestimated (i.e., under-generated) in the PA data will not be considered as important, and this distribution will only be falsely validated by the PO data that has been doubly used.

Therefore, we would like to explore how the effect of sampling bias (mispecification of the PP) can effect the PA generation. To do so, we will conduct the following experiments:
  
  Assume a known distribution for $\theta$, p(\theta). We can then generate PA data in two ways:
  1. Generate PA $p(y|D)$ data by repeatedly drawing from this distribution and simulating from the conditional of y. Here we also generate PO data
by simulating from the draws of $\theta$ as the parameters induce both PO and PA processes.
2. Generate PO data by drawing from $p(\theta)$. We then fit the PP, draw from the posterior $p(\theta|y_0)$ and then simulate the conditional of $y$.

We then fit the SO model and see the extent to which we can recover the true occupancy pattern given varying amounts of sampling bias. 
```{r stan-models}
poisson_process <- '
    data{
      int<lower = 1> N;
      int<lower=0> k_i; // number of predictors for intensity
      int<lower=0> k_b; // number of predictors for bias
      matrix[N, k_i] X; // predictor matrix for intensity
      matrix[N, k_b] Z; // predictor matrix for bias
      array[N] int y;
    }
    parameters{
      real alpha;
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
    }
    model{
      //priors
      target += normal_lpdf(alpha | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(delta | 0,10);

      // likelihood
      // log parameterization so we dont have to exponentiate coefs before passing into poisson distribution
      // https://mc-stan.org/docs/functions-reference/poisson-distribution-log-parameterization.html
      target += poisson_log_lpmf(y | alpha + X * beta + gamma + Z * delta);
    }
    generated quantities{
      array[N] int y_rep;
      y_rep = poisson_log_rng(alpha + X * beta + gamma + Z * delta);
    }
'

poisson_process_site_occupancy <- '
    data{
      int<lower = 1> n_pa_sites;
      int<lower = 1> n_po_sites;
      int<lower=0> k_i;
      int<lower=0> k_b;
      int<lower=0, upper=1> model_diag;
      array[n_pa_sites] int Y;
      array[n_pa_sites] int n_surveys;
      matrix[n_pa_sites, k_i] X;
      array[n_po_sites] int PO;
      matrix[n_po_sites, k_i] X_po; // predictor matrix for intensity
      matrix[n_po_sites, k_b] Z_po; // predictor matrix for bias
    }
    parameters{
      real alpha;           // intercepts
      real gamma;
      vector[k_i] beta;       // vector of params for intensity
      vector[k_b] delta;      // vector of params for bias
      real<lower = 0, upper = 1> p;
    }
    model{
      // priors
      target += normal_lpdf(alpha |   0,10);
      target += normal_lpdf(gamma | 0,10);
      target += normal_lpdf(beta | 0,10);
      target += normal_lpdf(delta | 0,10);
      vector[n_pa_sites] g_theta;
      target += poisson_log_lpmf(PO | alpha + X_po*beta + gamma + Z_po*delta);
      g_theta = 1 - exp(-exp(alpha + X*beta));
      for (i in 1:n_pa_sites) {
        if (Y[i] > 0){
          target += log(g_theta[i]*choose(n_surveys[i], Y[i])*(p^Y[i])*(1-p)^(n_surveys[i]-Y[i]));
        }
        else{
          // Compute mixture of no detection and no occupancy
          target += log(g_theta[i]*(1-p)^(n_surveys[i]) + (1 - g_theta[i]));
        }
      }
    }
    generated quantities{
      // Required generated quantities
      // We can generate over all sites, instead of just those being used for PA
      vector[n_po_sites] g_theta_gen;
      array[n_po_sites] int occ_gen; 
      // Note that we do not include gamma here because the 
      // probability of occupancy is just P(N(C_i)>0) = G(alpha + beta X)
      g_theta_gen = 1 - exp(-exp(alpha + X_po*beta));
      // Posterior predictive distribution for occupancy
      occ_gen = bernoulli_rng(g_theta_gen);
    }
'
```
First set hyperparameters used throughout the data generation and model fitting:
  ```{r set-vars}
n <- 5
k <- 20
sites <- k^2
area_D <- 400
centroid <- c(10,4)
centroid_2 <- c(20,4)
exp_args <- list(bias_func="exponential")
link_func <- "cloglog"
data_reps <- 10
p_detection <- 0.2
```
Then generate the simulated sampling surface data
```{r surface-data}
get_bias_surface_misspecified <- function(sampling_grid, centroid, strength){
  x <- seq(1:k)
  y <- seq(1:k)
  # Generate all possible coordinate points
  grid_points <- expand.grid(x, y)
  # Then compute distance from centroid to every other location
  r <- apply(grid_points, 1, function(x, center_coord){sqrt((center_coord[1] - x[1])^2 + (center_coord[2] - x[2])^2)}, center_coord=centroid)
  # Then create z covariate
  x <- exp(-strength*((r)/5))
  x_1 <- qnorm(0.98*x + .01)
  sampling_grid$aux_e <- x_1
  return(sampling_grid)
}

sampling_surface <- get_sampling_surface_donut(k)
sampling_surface <- get_bias_surface_exponential(sampling_surface, centroid)
# Strength of 0.5
sampling_surface <- get_bias_surface_misspecified(sampling_surface, centroid_2, 0.5)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_x)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="X") +
  ggtitle("Initial sampling surface, X covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_z)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Z") +
  ggtitle("Initial sampling surface, Z covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=aux_e)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="E") +
  ggtitle("Initial sampling surface, E covariate") +
  theme(text=element_text(size=5), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

```
Define functions for data generation. The PA functions take either a distribution for $\theta$ or pre-computed prior values.
```{r data-gen-func}
# This function is a little different than in the optimal design code as I am using multiple prior draws
# of params instead of fixed values
generate_pa <- function(data_reps, surface_data, p_0, prior, n, sites, link){
  # Add a column of 1s to the auxiliary data
  a_intercept <- rep(1, nrow(surface_data))
  surface_data <- cbind(a_intercept, surface_data)
  occupancy_maps <- matrix(nrow=data_reps, ncol=sites)
  Y_detection <- matrix(nrow=data_reps, ncol=sites)
  theta_reps <- matrix(nrow=data_reps, ncol=sites)
  print(prior[1, ])
  print(as.matrix(prior[1, ]))
  print(matrix(surface_data))
  
  for (r in 1:data_reps){
    param_vec <- as.matrix(prior[r,])
    if (link=="cloglog"){
      g_theta <- 1 - exp(-exp(as.matrix(surface_data)%*%matrix(param_vec)))
    }else{
      stop("Link not available")
    }
    stopifnot(sites==length(g_theta))
    site_presence <- rbinom(sites, 1, g_theta)
    theta_reps[r, ] <- g_theta
    occupancy_maps[r, ] <- site_presence
    Y_detection[r, ] <- rbinom(sites, n, p_0*site_presence)
  } 
  return(list(occupancy=occupancy_maps, Y=Y_detection, theta=theta_reps))
}

generate_ppp_data_misspecified <- function(surface_data, params, n_sites, data_reps, corr_matrix, gp_bool, area_D){
  # Generate some random covariate data per site on a grid that will be used to model lambda and b
  # There will be r (data_reps) datasets replicated, for use with the exchange 
  # algorithm/monte carlo integration and to maintain efficiency in the parallel 
  # function computation of V
  lambda <- (area_D / n_sites) * exp(params$alpha + params$beta*surface_data$aux_x)
  eta <- exp(params$gamma + params$delta*surface_data$aux_z + params$epsilon*surface_data$aux_e)
  b <- eta / (1 + eta)
  if (gp_bool==T){
    gp <- exp(rnorm(n_sites, 0, corr_matrix))
    Y_counts <- rpois(n_sites, lambda * b * gp)
  }
  else{
    Y_counts <- rpois(n_sites, lambda * b)
  }
  # Generate coordinate points within each site for each observation.
  Y_coords <- generate_coords(surface_data, Y_counts)
  # Create matrix of data reps, though for po data we just copy everything,
  # as opposed to random generation for each rep.
  Y_po <- matrix(rep(Y_counts, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  lambdas <- matrix(rep(lambda, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  biases <- matrix(rep(b, data_reps), nrow=data_reps, ncol=n_sites, byrow=T)
  return(list(Y=Y_po, lambda=lambdas, bias=biases, Y_coords=Y_coords))
}

```
There are two data generating procedures that we will compare here: 
  
  First, draw from the prior $p(\theta)$. Then, using those draws, generate PA and PO data accordingly, and fit the SO model with a PP prior.

Second, draw from the prior $p(\theta_0)$. Then, generate PO data, fit the PP model, and use these estimates to generate PA data. Then fit the SO model with a PP prior, again using the generated PO data. 

For generating (one draw of) the PO data, we will need a fixed value of $\theta$. This will reflect the effect of the parameters on the data generating process for that particular data collection period. We can use the mean of the true distribution, or a draw, but it doesn't matter exactly what fixed value we use because the draws for the PA and PO data, regardless of the experiment, are coming from the same distribution. In the PP model, we take that uncertainty into account by incorporating a prior. The important thing is that, given the PO data, we see the extent to which it can bias the PA data generation and hence the SO model fitting, in comparison to a case in which the PA and PO data generating processes are not tied together explicitly through the PO and only through the parameters.

We run the first experiment now:
```{r oracle-data-gen}
if (over_prior == T){
  # Draw from priors
  alpha_draws <- rnorm(data_reps, -2, 0.5)
  beta_draws <- rnorm(data_reps, 1, 0.25)
  pa_param_draws <- cbind(alpha_draws, beta_draws)
}else{
  # Or just set to the mean and use as fixed
  alpha <- -2
  beta <- 1
  alpha_draws <- rep(alpha, data_reps)
  beta_draws <- rep(beta, data_reps)
  pa_param_draws <- cbind(alpha_draws, beta_draws)
}

# Use mean as fixed values for PO params 
gamma <- 0.1
delta <- 2
epsilon <- 1.5
po_param_vec <- list(alpha=alpha, beta=beta, gamma=gamma, delta=delta, epsilon=1.5)

# Using pre-defined funcky for PO
po_gen <- generate_ppp_data_misspecified(sampling_surface, po_param_vec, sites, data_reps, corr_matrix=NA, gp_bool=FALSE, area_D)
# And the function defined above for the PA data
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pa_param_draws, n, sites, link_func)

# Check the generated_data
text_size <- 10
# Plot PA data
for (d in 1:data_reps){
  subtitle <- paste("alpha=", alpha_draws[d],", beta=", beta_draws[d], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}

# Then the PO data. Only 1 set of repeated po data. The PO data is all the same over the rows (repeated for efficiency in OD)
p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$lambda[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, name="Lambda") +
  ggtitle("Generated intensity per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

p <- ggplot(sampling_surface, aes(x, y, fill=po_gen$bias[data_reps,])) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE, "Bias") +
  ggtitle("Generated bias per site") +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)

thinned_intensity <- po_gen$lambda[data_reps,]*po_gen$bias[data_reps,]
p <- ggplot() +
  geom_tile(sampling_surface, mapping=aes(x, y, fill=thinned_intensity, width=1, height=1), alpha=.6) + 
  scale_fill_viridis(discrete=FALSE, name="L*b") +
  ggtitle("Generated PP thinning per site, \nincluding observed PO individuals") +
  geom_point(data=po_gen$Y_coords, mapping=aes(x=x, y=y), size=2, col="white") +
  theme(panel.grid.minor = element_line(colour="white")) +
  scale_y_continuous(breaks = seq(0, 20, 1)) +
  scale_x_continuous(breaks = seq(0, 20, 1)) +
  theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
  coord_fixed()
print(p)
```
First compile the model for both experiments
```{r stan-compile}
model_path <- "pp_so_for_sim.stan"
model_string <- poisson_process_site_occupancy
write(model_string, model_path)
model_so <- cmdstan_model(model_path) 
```

One question is what effect does sampling $m$ subset of sites versus using the full set of sites in the the SO model in the context of this experiment.

In the model below, it is fit to to just the first replication of PA data here. This means that we are not taking the 
uncertainty of Y into account in these experiments. We just consider a fixed dataset for fixed parameters. 
```{r oracle-stan-fit}
# Using all sites for now
m <- sites
#site_idx <- sample(which(pa_gen$Y[10,]>0), m, replace=F)
#site_idx <- sample(1:sites, m, replace=F)
#site_idx
covars_sampled_sites <- sampling_surface$aux_x
dim(as.matrix(covars_sampled_sites))
#covars_sampled_sites
# The PA counts are those generated from the PP parameters above.
# For the sake of illustration we pick 1 dataset
#selected_counts <- pa_gen$Y[10,site_idx]
#selected_counts
# n is for number of visits
n_surveys <- rep(n, sites)
k_i <- 1
k_b <- 1

data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,], X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
# Using the earlier compiled model, fit to PO+PA data
fit_so_1 <- model_so$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500) 
```
```{r oracle-diag}
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_1$summary(all_params)
posterior <- fit_so_1$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_so_1$draws(), pars=all_params)
mcmc_hist(fit_so_1$draws(), pars = all_params)
mcmc_pairs(fit_so_1$draws(), pars=all_params)
mcmc_scatter(fit_so_1$draws(), pars=c('alpha', 'gamma'))

# Examine the ppd for y
generated_occ <- fit_so_1$draws("occ_gen", format="matrix")
generated_prob <- fit_so_1$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```

In the next experiment, we use the same generated PO dataset, using the same values of the parameters.
Use the same generated PO dataset. We then fit the PP model to this dataset, and use the estimates from this model to generate the PA data.
```{r po-est}
data_site_occ = list(N=sites, X=as.matrix(sampling_surface$aux_x), y=po_gen$Y[1,], Z=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b)
model_path <- "poisson_process_sim.stan"
model_string <- poisson_process
write(model_string, model_path)
model_pp <- cmdstan_model(model_path) 
fit_pp <- model_pp$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500) 
```

```{r pp-diag}
fit_pp$summary(all_params)
posterior <- fit_pp$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_pp$draws(), pars=all_params)
mcmc_hist(fit_pp$draws(), pars = all_params)
mcmc_pairs(fit_pp$draws(), pars=all_params)
mcmc_scatter(fit_pp$draws(), pars=c('alpha', 'gamma'))

generated_y <- fit_pp$draws("y_rep", format="matrix")
yrep_mean <- colMeans(generated_y)
ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, yrep_mean=yrep_mean)

p <- ggplot(ppd_df, aes(x, y, fill=yrep_mean)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD Counts per site")
print(p)
```

```{r pa-gen-with-pp}
pp_posterior_draws <- fit_pp$draws(c("alpha", "beta[1]"), format="matrix")
#fit_pp$draws(["gamma"])
colnames(pp_posterior_draws)
dim(pp_posterior_draws)
# How do these estimates compare to the true means for alpha and beta used to generate PO?
print(colMeans(pp_posterior_draws))
# Get 1000 samples from the posterior
if (over_prior==TRUE){
  p_index <- sample(c(1:data_reps), size = data_reps, replace = T)
  pp_posterior <- pp_posterior_draws[p_index, ]
} else{
  pp_posterior <- matrix(rep(colMeans(pp_posterior_draws), data_reps), nrow=data_reps, byrow = T)
}
stopifnot(dim(pa_param_draws)==dim(pp_posterior))

# Generate a new set of pa data reps
pa_gen <- generate_pa(data_reps, sampling_surface[c("aux_x")], p_detection, pp_posterior, n, sites, link_func)

# Check the generated_data over the iterations
text_size <- 10
for (d in 1:data_reps){
  subtitle <- paste("alpha=", pp_posterior[d, 1],", beta=", pp_posterior[d, 2], sep="")
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$Y[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Counts") +
    ggtitle("Generated counts per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$occupancy[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Occupancy") +
    ggtitle("Generated occupancy per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
  
  p <- ggplot(sampling_surface, aes(x, y, fill=pa_gen$theta[d,])) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE, name="Theta") +
    ggtitle("generated occupancy probability per site", subtitle=subtitle) +
    theme(text=element_text(size=text_size), legend.key.size = unit(0.25, 'cm')) +
    coord_fixed()
  print(p)
}
```
Using the newly generated PA data, we fit another SO model
```{r estimated-po-stan-fit}
data_site_occ = list(n_surveys=n_surveys, n_pa_sites=m, n_po_sites=sites, X=as.matrix(covars_sampled_sites), Y=pa_gen$Y[1, ], PO=po_gen$Y[1,],
                         X_po=as.matrix(sampling_surface$aux_x), Z_po=as.matrix(sampling_surface$aux_z), k_i=k_i, k_b=k_b, model_diag=0)
# Using the earlier compiled model, fit to PO+PA data
fit_so_2 <- model_so$sample(data=data_site_occ, seed=13, chains=3, iter_sampling=2000, iter_warmup=500) 
```



```{r estimated-diag}
all_params <- c("alpha", "beta[1]", "gamma", "delta[1]")
params_intercept <- c("alpha", "gamma")
fit_so_2$summary(all_params)
posterior <- fit_so_2$draws(all_params)

color_scheme_set("mix-blue-pink")
p_trace <- mcmc_trace(posterior,
                      facet_args = list(nrow = 2, labeller = label_parsed))
print(p_trace + facet_text(size = 15))

plot_title <- ggtitle(paste("Posterior distributions, with means and 90% interval"))
for (aux_param in all_params){
  mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars=c(aux_param)) + plot_title
}

plot_title <- ggtitle(paste("Posterior distributions of detection probability, mean and 90% interval"))
p_post <- mcmc_areas(posterior,  prob = 0.8, point_est="mean", pars = params_intercept) + plot_title
print(p_post)

mcmc_intervals(fit_so_2$draws(), pars=all_params)
mcmc_hist(fit_so_2$draws(), pars = all_params)
mcmc_pairs(fit_so_2$draws(), pars=all_params)
mcmc_scatter(fit_so_2$draws(), pars=c('alpha', 'gamma'))

# Examine the ppd for y
generated_occ <- fit_so_2$draws("occ_gen", format="matrix")
generated_prob <- fit_so_2$draws("g_theta_gen", format="matrix")
dim(generated_occ)
occ_means <- colMeans(generated_occ)
get_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
occ_mode <- apply(generated_occ, MARGIN=2, FUN="get_mode")
length(occ_mode)
length(occ_means)
prob_means <- colMeans(generated_prob)

ppd_df <- data.frame(x=sampling_surface$x, y=sampling_surface$y, occ_mode=occ_mode, occ_rep=occ_means, theta_rep=prob_means)
head(ppd_df)

# Plot generated ys and thetas on the sampling grid
p <- ggplot(ppd_df, aes(x, y, fill=occ_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy per site")
print(p)

# theta_rep is already in prob space
p <- ggplot(ppd_df, aes(x, y, fill=theta_rep)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy probability per site")
print(p)

p <- ggplot(ppd_df, aes(x, y, fill=occ_mode)) + 
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    ggtitle("PPD occupancy mode per site")
print(p)
```

```{r compare}
print(fit_so_1$summary(all_params))
print(fit_so_2$summary(all_params))
```
